---
title: "R Notebook"
output: html_notebook
---

```{r setup}
local <- TRUE

wd <- ""

if (!local) {
  # Cluster environment
  wd <- "~/Documents/Autocorrelation"
  # Base directory for the data set we will create
  dat_dir <- "/project_cephfs/3022017.02/projects/lorkno/data"
} else {
  wd <- "~/HPC/Documents/Autocorrelation"
  # Base directory for the data set we will create
  dat_dir <- "~/HPC_project/data"
}

model_dir <- file.path(dat_dir, "complete_data_models")

# For some reason this sets the pwd only for the current scope, which means it 
# does not affect anything outside an if block if you set it there.
# So that's why it's here instead of above.
setwd(wd) 

source("src/load_dependencies.R")
```

# Global GLM for TDIA

```{r}
load(file.path(dat_dir, "models_tdia.rda")) # Legacy code
# models_tdia <- readRDS(file.path(model_dir, "models_tdia.rds"))
```

```{r}
pooled <- pool(imputed_models)
pooled
```

```{r}
summary(pooled)
```

Nothing significant...

# Individual GLMs

## All covariates

```{r}
load(file.path(dat_dir, "imp_multi.rda"))
load(file.path(dat_dir, "dat_reg.rda"))
```

```{r}
mean_acc <- mean(dat_reg$mean_accuracy, na.rm = TRUE)

# 35 rows, with a sub-data-frame (imputed data set) in column `data`
all_imps <- mice::complete(imp_multi, "long") %>%
  mutate(mean_accuracy = mean_accuracy + mean_acc) %>%
  select(!NSSIyn) %>% # Sometimes only has 1 level, glm can't handle that
  group_by(.imp, subject) %>%
  nest()

all_imps
```

```{r warning=FALSE}
n_sub <- n_distinct(all_imps$subject)
imp_models <- vector("list", n_sub)
imp_pools <- vector("list", n_sub)
imp_pool_summaries <- vector("list", n_sub)

for (i in 1:n_sub) {
  print(str_glue("Subject {i} out of {n_sub}"))
  
  sub_imp <- all_imps %>%
    filter(subject == i)
  
  imp_models[[i]] <- sub_imp %>%
    rowwise() %>%
    mutate(model = list(glm(mean_accuracy ~ . - .id, 
                            data = data, 
                            family = "binomial")))
  
  imp_pools[[i]] <- pool(imp_models[[i]]$model)
  imp_pool_summaries[[i]] <- summary(imp_pools[[i]])
}
```

```{r}
for (i in 1:n_sub) {
  if (any(imp_pool_summaries[[i]]$p.value < 0.05, na.rm = TRUE)) {
    print(str_glue("Subject {subjects[i]}"))
    print(imp_pool_summaries[[i]])
  }
}
```

## numdrinks_yest

```{r}
load(file.path(dat_dir, "imp_multi.rda"))
load(file.path(dat_dir, "dat_reg.rda"))
```

```{r}
mean_numdrinks_yest <- mean(dat_reg$numdrinks_yest, na.rm = TRUE)
```

```{r}
imp_models_ndrinks <- mice::complete(imp_multi, "long") %>%
  mutate(numdrinks_yest = round(numdrinks_yest + mean_numdrinks_yest)) %>%
  group_by(.imp) %>%
  nest() %>%
  rowwise() %>%
  mutate(model = list(glmer(numdrinks_yest ~ mean_accuracy + (1 | subject), 
                            data = data,
                            family = "poisson")))

imp_models_ndrinks
```

```{r}
pooled_ndrinks <- pool(imp_models_ndrinks$model)
pooled_ndrinks
```

```{r}
summary(pooled_ndrinks)
```

# Pooling with more BiAffect measures 

Global centering and scaling

## Without autocorrelation

```{r}
# About 7.8 GB, might be quicker to generate locally
load(file.path(dat_dir, "imputed_models_tdia_30_blocked_centered_scaled.rda"))
```

```{r}
all_pools <- imputed_models %>%
  summarize(across(menstrualbleeding:n_stressors, ~ list(pool(.x))))

all_pools
```

```{r}
for (i in 1:ncol(all_pools)) {
  summ <- summary(all_pools[[i]][[1]])
  
  if (any(summ$p.value <= 0.05)) {
    print(colnames(all_pools)[i])
    print(summ)
    print("======================")
  }
}
```

## With autocorrelation

```{r}
load(file.path(dat_dir, "pooling", "pooled_all_ema_autocor.rda"))
```

```{r}
for (i in 1:length(all_pools)) {
  summs[[i]] <- summary(all_pools[[i]])

  if (any(summs[[i]]$p.value <= 0.05)) {
    print(ema_cols[i])
    print(summs[[i]])
    print("======================")
  }
}
```

# Affective instability with RMSSD - Collapse all

```{r}
load(file.path(dat_dir, "imp_multi_tdia_30_blocked.rda"))
load(file.path(dat_dir, "dat_reg_scales_tdia_30_blocked.rda"))
```

Note: Perhaps I should calculate a time-independent identification accuracy, instead of taking the mean of the time-dependent identification accuracy.

```{r}
cdat <- complete(imp_multi, 1)

rmssd <- function(x) sqrt(mean(diff(x)^2))

cdat %>%
  group_by(subject) %>%
  summarize(
    tdia = mean(mean_accuracy),
    across(menstrualbleeding:n_stressors, rmssd)
  )
```

## Predict TDIA

```{r}
mean_acc <- centers["mean_accuracy"]

imp_models_rmssd <- mice::complete(imp_multi, "long") %>%
  select(!NSSIyn) %>%
  group_by(.imp, subject) %>%
  summarize(
    tdia = mean(mean_accuracy + mean_acc),
    across(menstrualbleeding:n_stressors, rmssd)
  ) %>%
  group_by(.imp) %>%
  nest() %>%
  rowwise() %>%
  mutate(model = list(glm(tdia ~ SleepLNQuality,
                          data = data,
                          family = "binomial")))
```

```{r}
pooled_rmssd <- pool(imp_models_rmssd$model)
pooled_rmssd
```

```{r}
summary(pooled_rmssd)
```


## Predict EMA instability

```{r}
mean_acc <- centers["mean_accuracy"]
scale_acc <- scales["mean_accuracy"]

imp_rmssd <- mice::complete(imp_multi, "long") %>%
  select(!NSSIyn) %>%
  group_by(.imp, subject) %>%
  summarize(
    tdia = mean(mean_accuracy * scale_acc + mean_acc),
    across(menstrualbleeding:n_stressors, rmssd)
  )

imp_rmssd
```

```{r}
imp_ema_rmssd_models <- imp_rmssd %>%
  group_by(.imp) %>%
  summarize(
    across(menstrualbleeding:n_stressors, ~ list(lm(.x ~ tdia)))
  )
```

```{r}
imp_ema_rmssd_pooled <- imp_ema_rmssd_models %>%
  ungroup() %>%
  summarize(across(menstrualbleeding:n_stressors, ~ list(pool(.x))))

imp_ema_rmssd_pooled
```

```{r}
for (i in 1:ncol(imp_ema_rmssd_pooled)) {
  summ <- summary(imp_ema_rmssd_pooled[[i]][[1]])
  
  if (any(summ$p.value <= 0.05)) {
    print(colnames(imp_ema_rmssd_pooled)[i])
    print(summ)
    print("======================")
  }
}
```

### With more BiAffect measures

```{r}
load(file.path(dat_dir, "imp_multi_tdia_30_blocked.rda"))
load(file.path(dat_dir, "dat_reg_tdia_30_blocked.rda"))
```

```{r}
imp_comp <- mice::complete(imp_multi, "long") %>%
  select(!NSSIyn) %>%
  mutate(week = rep(week(dat_reg$date), times = max(.imp))) %>%
  group_by(.imp, subject) %>%
  summarize(
    tdia = mean(mean_accuracy),
    medianIKD = mean(medianIKD),
    percent95IKD = mean(percent95IKD),
    madIKD = mean(madIKD),
    autocorrectRate = mean(autocorrectRate),
    backspaceRate = mean(backspaceRate),
    meanTotalKeyPresses = mean(totalKeyPresses),
    active = mean(active),
    upright = mean(upright),
    bed = mean(bed),
    across(menstrualbleeding:n_stressors, rmssd)
  ) %>%
  ungroup()

imp_comp
```

```{r}
imp_ema_rmssd_models <- imp_comp %>%
  group_by(.imp) %>%
  summarize(across(menstrualbleeding:n_stressors,
                   ~ list(lm(.x ~ tdia +
                                 medianIKD +
                                 percent95IKD +
                                 madIKD +
                                 autocorrectRate +
                                 backspaceRate +
                                 meanTotalKeyPresses +
                                 active +
                                 upright +
                                 bed))))
```

```{r}
imp_ema_rmssd_pooled <- imp_ema_rmssd_models %>%
  ungroup() %>%
  summarize(across(menstrualbleeding:n_stressors, ~ list(pool(.x))))

imp_ema_rmssd_pooled
```

```{r}
for (i in 1:ncol(imp_ema_rmssd_pooled)) {
  summ <- summary(imp_ema_rmssd_pooled[[i]][[1]])
  
  if (any(summ$p.value <= 0.05)) {
    print(colnames(imp_ema_rmssd_pooled)[i])
    print(summ)
    print("======================")
  }
}
```

Nothing significant except the intercepts.

# Affective instability with RMSSD - Per week

```{r}
load(file.path(dat_dir, "imp_multi_tdia_30_blocked.rda"))
load(file.path(dat_dir, "dat_reg_tdia_30_blocked.rda"))
```

```{r}
imp_comp <- mice::complete(imp_multi, "long") %>%
  select(!NSSIyn) %>%
  mutate(week = rep(week(dat_reg$date), times = max(.imp))) %>%
  group_by(.imp, subject, week) %>%
  summarize(
    tdia = mean(mean_accuracy),
    medianIKD = mean(medianIKD),
    percent95IKD = mean(percent95IKD),
    madIKD = mean(madIKD),
    autocorrectRate = mean(autocorrectRate),
    backspaceRate = mean(backspaceRate),
    meanTotalKeyPresses = mean(totalKeyPresses),
    active = mean(active),
    upright = mean(upright),
    bed = mean(bed),
    across(menstrualbleeding:n_stressors, rmssd)
  ) %>%
  ungroup()

imp_comp
```

```{r}
imp_ema_weekly_rmssd_models <- imp_comp %>%
  group_by(.imp) %>%
  summarize(across(menstrualbleeding:n_stressors,
                   ~ list(lmer(.x ~ tdia +
                                 medianIKD +
                                 percent95IKD +
                                 madIKD +
                                 autocorrectRate +
                                 backspaceRate +
                                 meanTotalKeyPresses +
                                 active +
                                 upright +
                                 bed +
                                 (1 | subject)))))

imp_ema_weekly_rmssd_models
```

```{r}
imp_ema_weekly_rmssd_pooled <- imp_ema_weekly_rmssd_models %>%
  ungroup() %>%
  summarize(across(menstrualbleeding:n_stressors, ~ list(pool(.x))))

imp_ema_weekly_rmssd_pooled
```

```{r}
for (i in 1:ncol(imp_ema_weekly_rmssd_pooled)) {
  summ <- summary(imp_ema_weekly_rmssd_pooled[[i]][[1]])
  
  if (any(summ$p.value <= 0.05)) {
    print(colnames(imp_ema_weekly_rmssd_pooled)[i])
    print(summ)
    print("======================")
  }
}
```

Only significant intercepts.
