---
title: "R Notebook"
output: html_notebook
---

```{r setup}
local <- TRUE

wd <- ""

if (!local) {
  # Cluster environment
  wd <- "~/Documents/Autocorrelation"
  # Base directory for the data set we will create
  dat_dir <- "/project_cephfs/3022017.02/projects/lorkno/data"
} else {
  wd <- "/Volumes/home/preclineu/lorkno/Documents/Autocorrelation"
  # Base directory for the data set we will create
  dat_dir <- "~/HPC_project/data"
}

model_dir <- file.path(dat_dir, "complete_data_models")

# For some reason this sets the pwd only for the current scope, which means it 
# does not affect anything outside an if block if you set it there.
# So that's why it's here instead of above.
setwd(wd) 

source("src/load_dependencies.R")

# Because it's required by the Lancet
options(OutDec = "·")
```

# Set apart a portion of the data as validation set

```{r eval=FALSE}
set.seed(42)

# Minimal threshold for proportion of the data that goes towards validation
val_prop_thr <- 0.1

# All data sets have the same number of entries, so we'll just select the first
# one to determine our split
dat <- dats_f[[1]]

# Keep selecting subjects until val_prop is satisfied
subs <- unique(dat$subject)
subs_perm <- sample(subs)

val_prop <- 0

for (i in 1:length(subs_perm)) {
  sub <- subs_perm[i]
  
  val_prop <- val_prop + length(dat$subject[dat$subject == sub]) / nrow(dat)
  
  if (val_prop >= val_prop_thr) {
    val_subs <- subs_perm[1:i]
    break
  }
}

print(val_prop)
print(val_subs)
```

```{r eval=FALSE}
# # Validation data
# dats_v <- lapply(dats_f, function(dat) {
#   dat %>%
#     filter(subject %in% val_subs)
# })
# 
# # Training data
# dats_t <- lapply(dats_f, function(dat) {
#   dat %>%
#     filter(!(subject %in% val_subs))
# })

# For now, we use everything as training data
dats_t <- dats_f
```

# Data prep function

```{r}
prep_reg_dat <- function(dat_path, ica_path) {
  # Semi-contiguous or fragmented data
  dat_reg <- readRDS(dat_path)
  
  dat_bi <- dat_reg %>%
    filter(modality == "biaffect") %>%
    select(!c(study_start_date, treatment_start_date, modality)) %>%
    pivot_wider(names_from = variable, values_from = value) %>%
    select(!partition_idx)
  
  print(str_glue("Number of BiAffect cases: {nrow(dat_bi)}"))
  print(str_glue("\tNumber of subjects: {n_distinct(dat_bi$subject)}"))
  
  dat_sr <- dat_reg %>%
    filter(modality == "self_report") %>%
    select(!c(study_start_date, treatment_start_date, modality)) %>%
    pivot_wider(names_from = variable, values_from = value)
  
  print(str_glue("Number of self-report cases: {nrow(dat_sr)}"))
  print(str_glue("\tNumber of subjects: {n_distinct(dat_sr$subject)}"))
  
  # RDA file
  load(ica_path)
  ns_comp <- as.integer(names(icas))
  
  # Combine BiAffect data with self-report ICs
  dats <- lapply(icas, function(ica) {
    data.frame(ica$S) %>%
      add_column(subject = subs, date = dates) %>%
      inner_join(dat_bi, by = c("subject", "date")) %>%
      mutate(
        week = week(date),
        sub_week = paste(subject, week, sep = " - ") # For convenience
      ) %>%
      relocate(subject, date, week, sub_week, .before = X1)
  })
  
  # Trash weeks within subject with only one observation
  print(str_glue("Complete-case entries in combined data: {nrow(dats[[1]])}"))
  print(str_glue("\tNumber of subjects: {n_distinct(dats[[1]]$subject)}"))
  
  dats_f <- lapply(dats, function(dat) {
    dat %>%
      group_by(sub_week) %>%
      filter(n() > 1) %>%
      ungroup()
  })
  
  print(str_glue("Entries in data after filtering out weeks with one observation: {nrow(dats_f[[1]])}"))
  print(str_glue("\tNumber of subjects: {n_distinct(dats_f[[1]]$subject)}"))
  
  dats_c <- lapply(dats_f, function(dat) {
    dat %>%
      mutate(
        totalKeyPresses = log(totalKeyPresses),
        across(medianIKD:mean_accuracy, scale)
      )
  })
  
  list(
    dat_reg = dat_reg, 
    dat_bi = dat_bi, 
    dat_sr = dat_sr, 
    icas = icas,
    subs = subs,
    dates = dates,
    dats_c = dats_c
  )
}
```

```{r}
ls <- prep_reg_dat(
  dat_path = file.path(dat_dir, "dat_reg_semi_contiguous.rds"),
  ica_path = file.path(dat_dir, "dat_reg_contig_icas.rda")
)

dat_reg_contig <- ls$dat_reg
dat_bi <- ls$dat_bi
dat_sr <- ls$dat_sr
icas <- ls$icas
subs <- ls$subs
dates <- ls$dates
dats_c <- ls$dats_c
```


# Examine IC behavior per grouping factor

Different people might experience a week differently. A good week for me could be a bad week for you, so it seems warranted to model a random effect of week within subject. We thought about including month as well, but that might be highly collinear with week. Perhaps we could include a random effect of menstrual cycle as well, but that might saturate the model - we'll check later whether this is possible. All in all, we hypothesise the following random effects:

* Subject.
* Week (within subject).

The first thing we have to do is check whether this grouping structure is justified. We start by plotting the data.

Intermezzo: How correlated are active, upright, and bed?

```{r}
ggplot(dats_c[[1]]) +
  geom_point(aes(active, bed), color = "steelblue") +
  geom_point(aes(active, upright), color = "darkred")

cor(dats_c[[1]] %>% select(active, bed, upright))
```

Let's throw out bed, as it's very correlated with upright and active.

## 5 components

```{r}
dat_c1 <- dats_c[[1]]
```

Let's check if a random effect of subject is necessary:

```{r fig.height=10}
for (ic in str_subset(colnames(dat_c1), "^X\\d+$")) {
  g <- ggplot(dat_c1, aes(.data[[ic]], subject)) +
    geom_point() +
    labs(title = ic)
  
  print(g)
}
```

There is considerable between-subject variability. Do we see the same if we condition on a fixed effect?

```{r fig.height=10}
for (ic in str_subset(colnames(dat_c1), "^X\\d+$")) {
  g <- ggplot(dat_c1, aes(medianIKD, .data[[ic]])) +
    geom_point() +
    facet_wrap(~ subject, scales = "free_x") +
    labs(title = ic)
  
  print(g)
}
```

We do. Technically, we would need to check all fixed effects, but that would be a bit too much. I think we can be fairly certain random subject effects are necessary. If not, we can always go back.

Let's check for week nested within subject:

```{r fig.height=10}
for (sub in unique(dat_c1$subject)) {
  g <- ggplot(dat_c1 %>%
                filter(subject == sub) %>%
                mutate(sub_week = paste(subject, week(date), sep = " - ")), 
              aes(X1, sub_week)) +
    geom_point() +
    labs(title = sub)
  
  print(g)
}
```

It's a bit hard to see, but we do see different effects of week within subject (compare e.g. subject 3029, week 36, to subject 3024, week 36).

Now we will double check the necessity for the hypothesised random effects with `lmList`.

```{r}
m5.sub_list <- lme4::lmList(X1 ~ medianIKD +
                             percent95IKD +
                             madIKD +
                             autocorrectRate +
                             backspaceRate +
                             totalKeyPresses +
                             active +
                             upright | 
                              subject,
                            data = dat_c1)

m5.sub_list
```

```{r fig.height=10}
plot(confint(m5.sub_list))
```

We cannot make interval plots for `subject/week`, as we always have more predictors than observations (a maximum of 7 per week). We therefore do it for `week` instead.

```{r}
m5.week_list <- lme4::lmList(X1 ~ mean_accuracy + 
                               medianIKD +
                               percent95IKD +
                               madIKD +
                               autocorrectRate +
                               backspaceRate +
                               totalKeyPresses +
                               active +
                               upright | 
                                week,
                              data = dat_c1)

m5.week_list
```

```{r fig.height=10}
plot(confint(m5.week_list))
```

## 10 components

```{r}
dat_c2 <- dats_c[[2]]
```

Let's check if a random effect of subject is necessary:

```{r fig.height=10}
for (ic in str_subset(colnames(dat_c2), "^X\\d+$")) {
  g <- ggplot(dat_c2, aes(.data[[ic]], subject)) +
    geom_point() +
    labs(title = ic)
  
  print(g)
}
```

It is.

# Mixed-effects models 

## 5 components

```{r}
dat_c1 <- dats_c[[1]]
```

```{r}
m5.X1 <- lmer(X1 ~ medianIKD +
               percent95IKD +
               madIKD +
               autocorrectRate +
               backspaceRate +
               totalKeyPresses +
               active +
               upright +
               (1 | subject / week),
             data = dat_c1)

m5.X1
```

```{r}
summary(m5.X1)
```

### Evaluate residuals

```{r}
qqnorm(resid(m5.X1))
qqline(resid(m5.X1))
```

```{r fig.height=10}
df <- data.frame(
  subject = dat_c1$subject, 
  week = dat_c1$week, 
  residuals = resid(m5.X1)
)

ggplot(df, aes(residuals, subject)) +
  geom_point()
```

There are a couple of subjects with rather large (negative) residuals. Otherwise, they are distributed quite normally.

```{r}
plot(m5.X1)
```

No heteroskedasticity.

```{r}
plot(m5.X1, subject ~ resid(.))
```

Residuals are roughly centered at 0 for all subjects, but their variance differs per subject. Might be because some subjects have relatively few observations.

```{r fig.height=10}
plot(m5.X1, subject ~ resid(.) | week, abline = 0)
```

Not all week-specific estimates are centered around 0. Perhaps we need to allow different variances per subject or week within subject.

```{r eval=FALSE}
m5.X1.var <- lme(X1 ~ medianIKD +
                   percent95IKD +
                   madIKD +
                   autocorrectRate +
                   backspaceRate +
                   totalKeyPresses +
                   active +
                   upright,
                 random = ~ 1 | subject / week,
                 data = dat_c1,
                 weights = varIdent(form = ~ 1 | subject * week),
                 control = lmeControl(msMaxEval = 500))

m5.X1.var
```

Sadly, the model above does not converge. Not too unexpected, as it as asks for lots of different variance parameters.

### Evaluate random effects

```{r}
re_sub <- ranef(m5.X1)$subject$`(Intercept)`
re_week <- ranef(m5.X1)$`week:subject`$`(Intercept)`

qqnorm(re_sub)
qqline(re_sub)
```

Hm, the tails of the random subject effects distribution are quite small.

```{r}
qqnorm(re_week)
qqline(re_week)
```

The distribution of the week within subject random effects is a bit skewed, but otherwise looks quite good.

### Miscellaneous evaluation

```{r fig.height=10}
plot(m5.X1, X1 ~ fitted(.) | subject, abline = c(0, 1))
```

Is week within subject a good addition to the model?

```{r}
m5.X1.1 <- lmer(X1 ~ medianIKD +
                 percent95IKD +
                 madIKD +
                 autocorrectRate +
                 backspaceRate +
                 totalKeyPresses +
                 active +
                 upright +
                 (1 | subject),
               data = dat_c1)

anova(m5.X1, m5.X1.1, refit = FALSE)
```

Week within subject seems to be a significant addition to the model fit.

Just a quick check to see if bed would have been a significant addition to the model:

```{r}
m5.X1.2 <- lmer(X1 ~ medianIKD +
                 percent95IKD +
                 madIKD +
                 autocorrectRate +
                 backspaceRate +
                 totalKeyPresses +
                 active +
                 upright +
                 bed +
                 (1 | subject / week),
               data = dat_c1)

anova(m5.X1, m5.X1.2, refit = TRUE)
```

It would not; perfect.

How about a random effect of month within subject?

```{r}
dat_c1_m <- dat_c1 %>%
  mutate(month = month(date))

m5.X1.1.1 <- lmer(X1 ~ medianIKD +
                   percent95IKD +
                   madIKD +
                   autocorrectRate +
                   backspaceRate +
                   totalKeyPresses +
                   active +
                   upright +
                   (1 | subject / week),
                 data = dat_c1_m)

m5.X1.1.2 <- lmer(X1 ~ medianIKD +
                   percent95IKD +
                   madIKD +
                   autocorrectRate +
                   backspaceRate +
                   totalKeyPresses +
                   active +
                   upright +
                   (1 | subject / week) +
                   (1 | subject:month),
                 data = dat_c1_m)

anova(m5.X1.1.1, m5.X1.1.2, refit = TRUE)
```

```{r}
m5.X1.1.1
```


```{r}
summary(m5.X1.1.2)
```

### Fixed-effect backwards fitting for IC 1

```{r eval=FALSE}
m5.X1.3 <- update(m5.X1, X1 ~ mean_accuracy + 
                           medianIKD +
                           percent95IKD +
                           madIKD +
                           autocorrectRate +
                           backspaceRate +
                           totalKeyPresses +
                           active +
                           (1 | subject / week))
anova(m5.X1, m5.X1.3)
```

```{r eval=FALSE}
m5.X1.4 <- update(m5.X1.3, X1 ~ mean_accuracy + 
                             medianIKD +
                             percent95IKD +
                             madIKD +
                             autocorrectRate +
                             backspaceRate +
                             active +
                             (1 | subject / week))

anova(m5.X1.3, m5.X1.4)
```

```{r eval=FALSE}
m5.X1.5 <- update(m5.X1.4, X1 ~ mean_accuracy + 
                             medianIKD +
                             percent95IKD +
                             madIKD +
                             autocorrectRate +
                             active +
                             (1 | subject / week))

anova(m5.X1.4, m5.X1.5)
```

```{r eval=FALSE}
m5.X1.6 <- update(m5.X1.5, X1 ~ mean_accuracy + 
                             medianIKD +
                             percent95IKD +
                             madIKD +
                             active +
                             (1 | subject / week))

anova(m5.X1.5, m5.X1.6)
```

```{r eval=FALSE}
m5.X1.7 <- update(m5.X1.6, X1 ~ mean_accuracy + 
                             medianIKD +
                             percent95IKD +
                             active +
                             (1 | subject / week))

anova(m5.X1.6, m5.X1.7)
```

```{r eval=FALSE}
m5.X1.8 <- update(m5.X1.7, X1 ~ mean_accuracy + 
                             medianIKD +
                             active +
                             (1 | subject / week))

anova(m5.X1.7, m5.X1.8)
```

```{r eval=FALSE}
m5.X1.9 <- update(m5.X1.8, X1 ~ mean_accuracy + 
                             active +
                             (1 | subject / week))

anova(m5.X1.8, m5.X1.9)
```

```{r eval=FALSE}
m5.X1.10 <- update(m5.X1.9, X1 ~ active +
                              (1 | subject / week))

anova(m5.X1.9, m5.X1.10)
```

```{r eval=FALSE}
m5.X1.11 <- update(m5.X1.10, X1 ~ (1 | subject / week))

anova(m5.X1.10, m5.X1.11)
```

```{r eval=FALSE}
m5.X1.12 <- update(m5.X1.10, X1 ~ active - 1 +
                              (1 | subject / week))

anova(m5.X1.10, m5.X1.12)
```

Final model:

```{r eval=FALSE}
summary(m5.X1.10)
```

```{r eval=FALSE}
m5.X1.10.nlme <- lme(X1 ~ active, dat_c1_m, ~ 1 | subject / week)
m5.X1.10.nlme
```

```{r paged.print=FALSE, eval=FALSE}
summ <- summary(m5.X1.10.nlme)
summ
```

```{r eval=FALSE}
summ$tTable
```


### Other ICs

Now rinse and repeat for the other ICs.

```{r}
m5.all <- dat_c1 %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(refit(m5.X1, .x, rename.response = TRUE))))

m5.all
```

```{r}
saveRDS(m5.all, file.path(dat_dir, "ic_5_models.rds"))
```


```{r}
lapply(m5.all, function(m) summary(m[[1]]))
```

```{r fig.height=10}
lapply(m5.all, function(m) {
  m <- m[[1]]
  resp <- formula(m)[[2]]
  plot(m, rlang::inject(!!resp ~ fitted(.) | subject), abline = c(0, 1))
})
```

```{r fig.height=8}
mix_df <- data.frame(icas$`5`$A) %>%
  rename_with(function(x) colnames(dat_sr)[3:ncol(dat_sr)]) %>%
  add_column(ic = 1:nrow(icas$`5`$A), .before = 1) %>%
  pivot_longer(!ic, names_to = "variable", values_to = "loading")

for (i in 1:ncol(m5.all)) {
  m <- m5.all[[i]][[1]]
  c <- capture.output(
    round(summary(m)$coefficients, 2)
  )
  # print(summary(m))
  
  g <- ggplot(mix_df %>% filter(ic == i), aes(loading, variable, fill = loading)) +
    geom_col() +
    scale_fill_gradient2(low = muted("blue"), mid = "white", high = muted("red"), 
                         name = "Loading") +
    geom_label(x = 0.5,
               y = "wishsleep",
               hjust = "left",
               vjust = "top",
               fill = "white",
               family = "mono",
               size = 4,
               label = paste(c, collapse = "\n")) +
    xlim(-0.5, 1.5) +
    xlab("Loading")
  
  print(g)
}
```

```{r fig.height=10}
mix_df <- data.frame(icas$`5`$A) %>%
  rename_with(function(x) colnames(dat_sr)[3:ncol(dat_sr)]) %>%
  add_column(ic = 1:nrow(icas$`5`$A), .before = 1) %>%
  pivot_longer(!ic, names_to = "variable", values_to = "loading")

summ_coeffs <- lapply(m5.all, function(m) {
  m <- m[[1]]
  paste(
    capture.output(
      round(summary(m)$coefficients[,-2], 2)
    ),
    collapse = "\n"
  )
})

coeff_df <- data.frame(ic = 1:5, text = as.vector(summ_coeffs, mode = "character"))

ggplot(mix_df, aes(loading, variable, fill = loading)) +
  geom_col() +
  scale_fill_gradient2(low = muted("blue"), mid = "white", high = muted("red"), 
                       name = "Loading") +
  geom_label(aes(label = text),
             x = 1,
             y = "wishsleep",
             hjust = "left",
             vjust = "top",
             fill = "white",
             family = "mono",
             size = 3.5,
             data = coeff_df) +
  xlim(-0.5, 4.5) +
  facet_wrap(~ ic, nrow = 2) +
  xlab("Loading")
```

```{r}
ggplot(dats_c[[1]], aes(active, X1)) +
  geom_point(alpha = 1)

ggplot(dats_c[[1]], aes(totalKeyPresses, X3)) +
  geom_point(alpha = 1)
```




```{r}
dat_c1_active <- dat_c1 %>%
  # add_column(fitted = fitted(m5.all[[1]][[1]])) %>%
  select(subject, date, X1, active) %>%
  pivot_longer(c(X1, active), "variable")

dat_c1_active
```

```{r fig.height=10}
ggplot(dat_c1_active, aes(date, value)) +
  geom_line(aes(color = variable)) +
  facet_wrap(~ subject, scales = "free_x")

ggplot(dat_c1_active %>% filter(subject == "3030"), aes(date, value)) +
  geom_line(aes(color = variable))
```


### Other ICs - NLME

NLME gives us p-values, necessary if we want to do multiple comparisons corrections.

```{r}
m5.X1.nlme <- lme(X1 ~ medianIKD +
                   percent95IKD +
                   madIKD +
                   autocorrectRate +
                   backspaceRate +
                   totalKeyPresses +
                   active +
                   upright,
                 random = ~ 1 | subject / week,
                 data = dat_c1)

m5.all.nlme <- dat_c1 %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(update(m5.X1.nlme, reformulate(".", cur_column())))))

m5.all.nlme
```

```{r paged.print=FALSE}
lapply(m5.all.nlme, function(m) summary(m[[1]]))
```

Get beta and significance values nicely formatted for copying into Word table.

```{r}
print_estimates <- function(models) {
  for (m in models) {
    summ <- summary(m[[1]])
    tab <- summ$tTable
    for (i in 1:nrow(tab)) {
      v1 <- signif(tab[i, 1], 2)
      v2 <- signif(tab[i, 5], 2)
      cat(str_glue("{v1}\t{v2}\n\n"))
    }
    
    cat("\n------------------------\n\n")
  }
}

print_estimates(m5.all.nlme)
```

Get Bonferroni-corrected p values.

```{r}
lapply(m5.all.nlme, function(m) {
  summ <- summary(m[[1]])
  tab <- summ$tTable
  for (i in 1:nrow(tab)) {
    cat(paste0(signif(min(tab[i,5] * 5 * 8, 1), 2), "\n"))
  }
  cat("\n")
})
```

```{r}
estimates_to_csv(file.path(dat_dir, "m5_nlme_p.csv"), 
                 m5.all.nlme, 
                 include_corrected = TRUE)
```


Check significance of random effects.

```{r}
lapply(m5.all.nlme, function(m) {
  m <- m[[1]]
  m_noweek <- update(m, random = ~ 1 | subject)
  
  anova(m, m_noweek)
})
```

Check for heteroskedasticity of residuals.

```{r}
lapply(m5.all.nlme, function(m) plot(m[[1]]))
```

Normality of residuals.

```{r}
lapply(m5.all.nlme, function(m) {
  m <- m[[1]]
  qqnorm(resid(m))
  qqline(resid(m))
})
```

Subject-level random effects.

```{r}
lapply(m5.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 1)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

Week-within-subjects-level random effects.

```{r}
lapply(m5.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 2)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

## 10 components

```{r}
dat_c2 <- dats_c[[2]]
```

```{r}
m10.X1 <- lmer(X1 ~ medianIKD +
                 percent95IKD +
                 madIKD +
                 autocorrectRate +
                 backspaceRate +
                 totalKeyPresses +
                 active +
                 upright +
                 (1 | subject / week),
               data = dat_c2)

m10.X1
```

```{r}
m10.all <- dat_c2 %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(refit(m10.X1, .x, rename.response = TRUE))))

m10.all
```

```{r}
lapply(m10.all, function(m) summary(m[[1]]))
```

```{r fig.height=8}
mix_df <- data.frame(icas$`10`$A) %>%
  rename_with(function(x) colnames(dat_sr)[3:ncol(dat_sr)]) %>%
  add_column(ic = 1:nrow(icas$`10`$A), .before = 1) %>%
  pivot_longer(!ic, names_to = "variable", values_to = "loading")

for (i in 1:ncol(m10.all)) {
  m <- m10.all[[i]][[1]]
  c <- capture.output(
    round(summary(m)$coefficients, 2)
  )
  # print(summary(m))
  
  g <- ggplot(mix_df %>% filter(ic == i), aes(loading, variable, fill = loading)) +
    geom_col() +
    scale_fill_gradient2(low = muted("blue"), mid = "white", high = muted("red"), 
                         name = "Loading") +
    geom_label(x = 0.5,
               y = "wishsleep",
               hjust = "left",
               vjust = "top",
               fill = "white",
               family = "mono",
               size = 4,
               label = paste(c, collapse = "\n")) +
    xlim(-0.5, 1.5) +
    xlab("Loading")
  
  print(g)
}
```

```{r}
summ_coeffs <- lapply(m10.all, function(m) {
  m <- m[[1]]
  paste(
    capture.output(
      round(summary(m)$coefficients[,-2], 2)
    ),
    collapse = "\n"
  )
})

coeff_df <- data.frame(ic = 1:10, text = as.vector(summ_coeffs, mode = "character"))

str(coeff_df)
```

```{r fig.height=10}
ggplot(mix_df, aes(loading, variable, fill = loading)) +
  geom_col() +
  scale_fill_gradient2(low = muted("blue"), mid = "white", high = muted("red"), 
                       name = "Loading") +
  geom_label(aes(label = text),
             x = 0.5,
             y = "wishsleep",
             hjust = "left",
             vjust = "top",
             fill = "white",
             family = "mono",
             size = 2.5,
             data = coeff_df) +
  xlim(-0.5, 4.5) +
  facet_wrap(~ ic, nrow = 2) +
  xlab("Loading")
```

### NLME

```{r}
m10.X1.nlme <- lme(X1 ~ medianIKD +
                     percent95IKD +
                     madIKD +
                     autocorrectRate +
                     backspaceRate +
                     totalKeyPresses +
                     active +
                     upright,
                   random = ~ 1 | subject / week,
                   data = dat_c2)

m10.all.nlme <- dat_c2 %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(update(m10.X1.nlme, reformulate(".", cur_column())))))

m10.all.nlme
```


```{r}
print_estimates_hor <- function(models) {
  for (m in models) {
    summ <- summary(m[[1]])
    tab <- summ$tTable
    
    v1s <- sapply(1:nrow(tab), function(i) signif(tab[i, 1], 2))
    v2s <- sapply(1:nrow(tab), function(i) signif(tab[i, 5], 2))
    
    str1 <- paste(rev(v1s), sep = " ")
    str2 <- paste(rev(v2s), sep = " ")
    
    cat(str1, "\n")
    cat(str2, "\n")
    
    cat("\n------------------------\n\n")
  }
}

sig_zeros <- function(x) str_replace(sprintf("%#.2g", x), "\\.", "·")

print_estimates_word <- function(models) {
  terms <- c("medianIKD", "percent95IKD", "madIKD", "autocorrectRate",
             "backspaceRate", "totalKeyPresses", "active", "upright")
  
  for (i in 1:length(terms)) {
    cat("\n", terms[i], "\n")
    
    term_ests <- lapply(models, function(m) {
      summ <- summary(m[[1]])
      tab <- summ$tTable
      
      beta <- sig_zeros(tab[terms[i], 1])
      p_val <- sig_zeros(tab[terms[i], 5])
      
      paste(beta, p_val, sep = "\n")
    })
    
    cat(paste(term_ests, collapse = "\n"), "\n")
  }
}

estimates_to_csv(file.path(dat_dir, "m10_nlme_p.csv"), m10.all.nlme, TRUE)
```

```{r}
lapply(m10.all.nlme, function(m) {
  summ <- summary(m[[1]])
  tab <- summ$tTable
  for (i in 1:nrow(tab)) {
    cat(paste0(rownames(tab)[i], ": ", signif(min(tab[i,5] * 10 * 8, 1), 2), "\n"))
  }
  cat("\n")
})
```


### Checking model assumptions

Check for heteroskedasticity of residuals.

```{r}
lapply(m10.all.nlme, function(m) plot(m[[1]]))
```

Normality of residuals.

```{r}
lapply(m10.all.nlme, function(m) {
  m <- m[[1]]
  qqnorm(resid(m))
  qqline(resid(m))
})
```

Subject-level random effects.

```{r}
lapply(m10.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 1)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

Week-within-subjects-level random effects.

```{r}
lapply(m10.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 2)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

## 20 components

```{r}
dat_c3 <- dats_c[[3]]
```

```{r}
m20.X1 <- lmer(X1 ~ medianIKD +
                 percent95IKD +
                 madIKD +
                 autocorrectRate +
                 backspaceRate +
                 totalKeyPresses +
                 active +
                 upright +
                 (1 | subject / week),
               data = dat_c3)

m20.X1
```

```{r}
m20.all <- dat_c3 %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(refit(m20.X1, .x, rename.response = TRUE))))

m20.all
```

```{r fig.height=8}
mix_df <- data.frame(icas$`20`$A) %>%
  rename_with(function(x) colnames(dat_sr)[3:ncol(dat_sr)]) %>%
  add_column(ic = 1:nrow(icas$`20`$A), .before = 1) %>%
  pivot_longer(!ic, names_to = "variable", values_to = "loading")

for (i in 1:ncol(m20.all)) {
  m <- m20.all[[i]][[1]]
  c <- capture.output(
    round(summary(m)$coefficients, 2)
  )
  # print(summary(m))
  
  g <- ggplot(mix_df %>% filter(ic == i), aes(loading, variable, fill = loading)) +
    geom_col() +
    scale_fill_gradient2(low = muted("blue"), mid = "white", high = muted("red"), 
                         name = "Loading") +
    geom_label(x = 0.5,
               y = "wishsleep",
               hjust = "left",
               vjust = "top",
               fill = "white",
               family = "mono",
               size = 4,
               label = paste(c, collapse = "\n")) +
    xlim(-0.5, 1.5) +
    xlab("Loading")
  
  print(g)
}
```

```{r fig.height=10}
mix_df <- data.frame(icas$`20`$A) %>%
  rename_with(function(x) colnames(dat_sr)[3:ncol(dat_sr)]) %>%
  add_column(ic = 1:nrow(icas$`20`$A), .before = 1) %>%
  pivot_longer(!ic, names_to = "variable", values_to = "loading")

summ_coeffs <- lapply(m20.all, function(m) {
  m <- m[[1]]
  ts <- round(summary(m)$coefficients[,3], 2)
  paste(
    paste(
      c("Int", "acc", "ikd", "95i", "mad", "aut", "bac", "key", "act", "upr"),
      ts
    ),
    collapse = "\n"
  )
})

coeff_df <- data.frame(ic = 1:20, text = as.vector(summ_coeffs, mode = "character"))

ggplot(mix_df, aes(loading, variable, fill = loading)) +
  geom_col() +
  scale_fill_gradient2(low = muted("blue"), mid = "white", high = muted("red"), 
                       name = "Loading") +
  geom_label(aes(label = text),
             x = 0.5,
             y = "wishsleep",
             hjust = "left",
             vjust = "top",
             fill = "white",
             family = "mono",
             size = 2.5,
             data = coeff_df) +
  xlim(-0.5, 2) +
  facet_wrap(~ ic, nrow = 2) +
  xlab("Loading")
```

### NLME

```{r}
m20.X1.nlme <- lme(X1 ~ medianIKD +
                     percent95IKD +
                     madIKD +
                     autocorrectRate +
                     backspaceRate +
                     totalKeyPresses +
                     active +
                     upright,
                   random = ~ 1 | subject / week,
                   data = dat_c3)

m20.all.nlme <- dat_c3 %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(update(m20.X1.nlme, reformulate(".", cur_column())))))

m20.all.nlme
```

```{r}
print_estimates_word(m20.all.nlme)
```

Bonferroni-corrected p-values.

```{r}
lapply(m20.all.nlme, function(m) {
  summ <- summary(m[[1]])
  tab <- summ$tTable
  for (i in 1:nrow(tab)) {
    cat(paste0(rownames(tab)[i], ": ", signif(min(tab[i,5] * 20 * 8, 1), 2), "\n"))
  }
  cat("\n")
})
```

### Checking model assumptions

Check for heteroskedasticity of residuals.

```{r}
lapply(m20.all.nlme, function(m) plot(m[[1]]))
```

Normality of residuals.

```{r}
lapply(m20.all.nlme, function(m) {
  m <- m[[1]]
  qqnorm(resid(m))
  qqline(resid(m))
})
```

Subject-level random effects.

```{r}
lapply(m20.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 1)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

Week-within-subjects-level random effects.

```{r}
lapply(m20.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 2)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

# Fragmented data

```{r}
ls <- prep_reg_dat(
  dat_path = file.path(dat_dir, "dat_reg_fragmented.rds"),
  ica_path = file.path(dat_dir, "dat_reg_frag_icas.rda")
)

dat_reg_frag <- ls$dat_reg
dat_bi <- ls$dat_bi
dat_sr <- ls$dat_sr
icas <- ls$icas
subs <- ls$subs
dates <- ls$dates
dats_c_frag <- ls$dats_c
```

```{r}
dat_c1_frag <- dats_c_frag[[1]]
```


## Examining need for random effects

Now we will double check the necessity for the hypothesised random effects with `lmList`.

```{r}
m5.frag.sub_list <- lme4::lmList(X1 ~ medianIKD +
                                   percent95IKD +
                                   madIKD +
                                   autocorrectRate +
                                   backspaceRate +
                                   totalKeyPresses +
                                   active +
                                   upright | 
                                   subject,
                                 data = dat_c1_frag)

m5.frag.sub_list
```

```{r fig.height=10}
plot(confint(m5.frag.sub_list))
```

We cannot make interval plots for `subject/week`, as we always have more predictors than observations (a maximum of 7 per week). We therefore do it for `week` instead.

```{r}
m5.frag.week_list <- lme4::lmList(X1 ~ mean_accuracy + 
                                    medianIKD +
                                    percent95IKD +
                                    madIKD +
                                    autocorrectRate +
                                    backspaceRate +
                                    totalKeyPresses +
                                    active +
                                    upright | 
                                    week,
                                  data = dat_c1_frag)

m5.frag.week_list
```

```{r fig.height=10}
plot(confint(m5.frag.week_list))
```

## 5 components

```{r}
m5.frag.X1 <- lmer(X1 ~ medianIKD +
                     percent95IKD +
                     madIKD +
                     autocorrectRate +
                     backspaceRate +
                     totalKeyPresses +
                     active +
                     upright +
                     (1 | subject / week),
                   data = dat_c1_frag)

m5.frag.X1
```

```{r}
summary(m5.frag.X1)
```

NLME gives us p-values, necessary if we want to do multiple comparisons corrections.

```{r}
m5.frag.X1.nlme <- lme(X1 ~ medianIKD +
                         percent95IKD +
                         madIKD +
                         autocorrectRate +
                         backspaceRate +
                         totalKeyPresses +
                         active +
                         upright,
                       random = ~ 1 | subject / week,
                       data = dat_c1_frag)

m5.frag.all.nlme <- dat_c1_frag %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(update(m5.frag.X1.nlme, 
                                 reformulate(".", cur_column())))))

m5.frag.all.nlme
```

```{r paged.print=FALSE}
lapply(m5.frag.all.nlme, function(m) summary(m[[1]]))
```

```{r}
p_df <- estimates_to_csv(file.path(dat_dir, "m5_nlme_p_frag.csv"), 
                         m5.frag.all.nlme, 
                         include_corrected = TRUE)

p_df
```

### Validation

Check for heteroskedasticity of residuals.

```{r}
lapply(m5.frag.all.nlme, function(m) plot(m[[1]]))
```

Lethargy model shows some structure in residuals. Let's plot actual values vs fitted values.

```{r}
fits <- fitted(m5.frag.all.nlme[[2]][[1]])
plot(fits, dat_c1_frag$X2)
abline(0, 1)
grid()
```

There is a rather large proportion of -1 values that are not captured very well by the model fit. These might correspond to the 0 ratings on the Likert scales.

```{r}
hist(dat_c1_frag$X2)
```

Plotting the active values vs raw X2 shows very little correlation:

```{r}
plot(dat_c1_frag$X2, dat_c1_frag$active)
```

Only when we remove the random effects from the fitted values, the correlation appears:

```{r}
plot(fitted(m5.frag.all.nlme[[2]][[1]], level = 0), dat_c1_frag$active)
```

This could be a point of criticism for this model.

Normality of residuals.

```{r}
lapply(m5.frag.all.nlme, function(m) {
  m <- m[[1]]
  qqnorm(resid(m))
  qqline(resid(m))
})
```

Subject-level random effects.

```{r}
lapply(m5.frag.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 1)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

Week-within-subjects-level random effects.

```{r}
lapply(m5.frag.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 2)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```


## 10 components

```{r}
dat_c2_frag <- dats_c_frag[[2]]

m10.frag.X1.nlme <- lme(X1 ~ medianIKD +
                           percent95IKD +
                           madIKD +
                           autocorrectRate +
                           backspaceRate +
                           totalKeyPresses +
                           active +
                           upright,
                         random = ~ 1 | subject / week,
                         data = dat_c2_frag)

m10.frag.all.nlme <- dat_c2_frag %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(update(m10.frag.X1.nlme, 
                                 reformulate(".", cur_column())))))

m10.frag.all.nlme
```

```{r}
estimates_to_csv(file.path(dat_dir, "m10_nlme_p_frag.csv"), 
                 m10.frag.all.nlme, 
                 include_corrected = TRUE,
                 transpose = TRUE)
```

### Validation

Check for heteroskedasticity of residuals.

```{r}
lapply(m10.frag.all.nlme, function(m) plot(m[[1]]))
```

Normality of residuals.

```{r}
lapply(m10.frag.all.nlme, function(m) {
  m <- m[[1]]
  qqnorm(resid(m))
  qqline(resid(m))
})
```

Subject-level random effects.

```{r}
lapply(m10.frag.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 1)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

Week-within-subjects-level random effects.

```{r}
lapply(m10.frag.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 2)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

## 20 components

```{r}
dat_c3_frag <- dats_c_frag[[3]]

m20.frag.X1.nlme <- lme(X1 ~ medianIKD +
                           percent95IKD +
                           madIKD +
                           autocorrectRate +
                           backspaceRate +
                           totalKeyPresses +
                           active +
                           upright,
                         random = ~ 1 | subject / week,
                         data = dat_c3_frag)

m20.frag.all.nlme <- dat_c3_frag %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(update(m20.frag.X1.nlme, 
                                 reformulate(".", cur_column())))))

m20.frag.all.nlme
```

```{r}
estimates_to_csv(file.path(dat_dir, "m20_nlme_p_frag.csv"), 
                 m20.frag.all.nlme, 
                 include_corrected = TRUE,
                 transpose = TRUE)
```

### Validation

Check for heteroskedasticity of residuals.

```{r}
lapply(m20.frag.all.nlme, function(m) plot(m[[1]]))
```

Normality of residuals.

```{r}
lapply(m20.frag.all.nlme, function(m) {
  m <- m[[1]]
  qqnorm(resid(m))
  qqline(resid(m))
})
```

Subject-level random effects.

```{r}
lapply(m20.frag.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 1)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```

Week-within-subjects-level random effects.

```{r}
lapply(m20.frag.all.nlme, function(m) {
  m <- m[[1]]
  refs <- ranef(m, level = 2)[[1]]
  qqnorm(refs)
  qqline(refs)
})
```
