---
title: "R Notebook"
output: html_notebook
---

```{r setup}
local <- TRUE

wd <- ""

if (!local) {
  # Cluster environment
  wd <- "~/Documents/Autocorrelation"
  # Base directory for the data set we will create
  dat_dir <- "/project_cephfs/3022017.02/projects/lorkno/data"
} else {
  wd <- "/Volumes/home/preclineu/lorkno/Documents/Autocorrelation"
  # Base directory for the data set we will create
  dat_dir <- "~/HPC_project/data"
}

model_dir <- file.path(dat_dir, "complete_data_models")

# For some reason this sets the pwd only for the current scope, which means it 
# does not affect anything outside an if block if you set it there.
# So that's why it's here instead of above.
setwd(wd) 

source("src/load_dependencies.R")
library(fastICA)

# Because it's required by the Lancet
options(OutDec = "·")
```

# Get/set seed

I would like to avoid setting the seed manually. I'll record my seed from today (14-06-2023) and use that in later runs of this script.

```{r}
# if(!exists(".Random.seed")) set.seed(NULL)
# seed <- .Random.seed
# saveRDS(seed, file.path(dat_dir, "sens_seed.rds"))

seed <- readRDS(file.path(dat_dir, "sens_seed.rds"))
```

# Prepare ICAs

For now, only do the 5-component solution.

```{r}
replicate_icas <- function(dat_path, n = 100, out_path = NULL) {
  set.seed(seed)
  
  ### Data prep
  
  dat_reg <- readRDS(dat_path)
  
  dat_reg_sr <- dat_reg %>%
    filter(modality == "self_report", !is.na(value)) %>%
    select(!c(study_start_date, treatment_start_date, modality)) %>%
    group_by(variable) %>%
    mutate(
      logged = case_when(
        min(value) == 0 ~ log(value + 1), # To prevent log(0) = NA values
        TRUE ~ log(value)
      ),
      scaled = scale(value)
    ) %>%
    ungroup()
  
  df <- dat_reg_sr %>%
    select(!c(value, scaled)) %>% 
    pivot_wider(names_from = variable, values_from = logged) %>%
    arrange(subject, date) 
  
  subs <- df$subject
  dates <- df$date
  
  df <- df %>%
    select(!c(subject, date))
  
  # t x q
  mat <- sapply(df, as.vector)
  
  means <- colMeans(mat)
  
  ### Running the ICAs
  
  icas <- replicate(
    n = n, 
    fastICA(mat, n.comp = 5, alg.typ = "parallel", fun = "logcosh"),
    simplify = FALSE
  )
  
  ls <- list(icas = icas, subs = subs, dates = dates)
  
  if (!is.null(out_path))
    saveRDS(ls, file = out_path)
  
  ls
}

ls <- replicate_icas(file.path(dat_dir, "dat_reg_semi_contiguous.rds"),
                     out_path = file.path(dat_dir, "icas_sens.rds"))

icas <- ls$icas
subs <- ls$subs
dates <- ls$dates
```

We use sens_helper.R to mark the general affect components and save the marked component indices to general_idx_sens.rds.

```{r}
general_idx <- readRDS(file.path(dat_dir, "general_idx_sens.rds"))
general_idx <- as.integer(general_idx)
```

# Run models

Prepare data.

```{r}
prep_reg_dat <- function(dat_path, icas, subs, dates, general_idx) {
  # Semi-contiguous data
  dat_reg <- readRDS(dat_path)
  
  dat_bi <- dat_reg %>%
    filter(modality == "biaffect") %>%
    select(!c(study_start_date, treatment_start_date, modality)) %>%
    pivot_wider(names_from = variable, values_from = value) %>%
    select(!partition_idx)
  
  # Select components that correspond to general affect components
  dats <- map2(icas, general_idx, function(ica, idx) {
    data.frame(X = ica$S[,idx]) %>%
      add_column(subject = subs, date = dates) %>%
      inner_join(dat_bi, by = c("subject", "date")) %>%
      mutate(
        week = week(date),
        sub_week = paste(subject, week, sep = " - ") # For convenience
      ) %>%
      relocate(subject, date, week, sub_week, .before = X)
  })
  
  dats_f <- lapply(dats, function(dat) {
    dat %>%
      group_by(sub_week) %>%
      filter(n() > 1) %>%
      ungroup()
  })
  
  dats_c <- lapply(dats_f, function(dat) {
    dat %>%
      mutate(
        totalKeyPresses = log(totalKeyPresses),
        across(medianIKD:mean_accuracy, scale)
      )
  })
  
  dats_c
}

dats_c <- prep_reg_dat(file.path(dat_dir, "dat_reg_semi_contiguous.rds"),
                       icas, subs, dates,
                       general_idx)
```

Create models.

```{r}
models <- lapply(dats_c, function(dat_c) {
  lme(X ~ medianIKD +
        percent95IKD +
        madIKD +
        autocorrectRate +
        backspaceRate +
        totalKeyPresses +
        active +
        upright,
      random = ~ 1 | subject / week,
      data = dat_c)
})
```

Extract p values.

```{r}
p_vals <- sapply(models, function(m) {
  summary(m)$tTable[,5]
})

p_val_df <- data.frame(t(p_vals)) %>%
  pivot_longer(everything(), names_to = "variable") %>%
  filter(variable != "X.Intercept.")

p_val_df
```


Plot p values.

```{r fig.height=10}
ggplot(p_val_df, aes(value)) +
  geom_histogram(bins = 60) +
  facet_wrap(~ variable, scales = "free_x", labeller = as_labeller(c(
    medianIKD = "Median IKD",
    percent95IKD = "95th percentile IKD",
    madIKD = "MAD IKD",
    autocorrectRate = "Autocorrect rate",
    backspaceRate = "Backspace rate",
    totalKeyPresses = "Total number of key presses",
    active = "Movement rate",
    upright = "Upright rate"
  ))) +
  ylab("Count") +
  xlab("p value") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, family = "serif"),
    strip.text = element_text(size = 14, family = "serif"),
    panel.border = element_rect(fill = NA, color = "grey")
  )
```

```{r fig.height=10}
ggplot(p_val_df, aes(value)) +
  geom_histogram(bins = 60) +
  facet_wrap(~ variable, scales = "free_x", labeller = as_labeller(c(
    medianIKD = "Median IKD",
    percent95IKD = "95th percentile IKD",
    madIKD = "MAD IKD",
    autocorrectRate = "Autocorrect rate",
    backspaceRate = "Backspace rate",
    totalKeyPresses = "Total number of key presses",
    active = "Movement rate",
    upright = "Upright rate"
  ))) +
  ylab("Count") +
  xlab("p value") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, family = "serif"),
    strip.text = element_text(size = 14, family = "serif"),
    panel.border = element_rect(fill = NA, color = "grey")
  )
```

```{r}
mean(p_val_df[p_val_df$variable == "active",]$value * 5 * 8 < 0.05)
```

# Fragmented data

```{r}
ls <- replicate_icas(file.path(dat_dir, "dat_reg_fragmented.rds"))
icas_frag <- ls$icas
subs_frag <- ls$subs
dates_frag <- ls$dates
```

```{r fig.height=10}
fancy_mix_fig(icas_frag[[1]])
```

Use sens_helper.R to get general_idx_frag.

```{r}
general_idx_frag <- readRDS(file.path(dat_dir, "general_idx_frag_sens2.rds"))
general_idx_frag <- as.integer(general_idx_frag)

dats_c_frag <- prep_reg_dat(file.path(dat_dir, "dat_reg_fragmented.rds"),
                       icas_frag, subs_frag, dates_frag,
                       general_idx_frag)

models_frag <- lapply(dats_c_frag, function(dat_c) {
  lme(X ~ medianIKD +
        percent95IKD +
        madIKD +
        autocorrectRate +
        backspaceRate +
        totalKeyPresses +
        active +
        upright,
      random = ~ 1 | subject / week,
      data = dat_c)
})

p_vals_frag <- sapply(models_frag, function(m) {
  summary(m)$tTable[,5]
})

p_val_df_frag <- data.frame(t(p_vals_frag)) %>%
  pivot_longer(everything(), names_to = "variable") %>%
  filter(variable != "X.Intercept.")
```

```{r fig.height=10}
ggplot(p_val_df_frag, aes(value)) +
  geom_histogram(bins = 60) +
  facet_wrap(~ variable, scales = "free_x", labeller = as_labeller(c(
    medianIKD = "Median IKD",
    percent95IKD = "95th percentile IKD",
    madIKD = "MAD IKD",
    autocorrectRate = "Autocorrect rate",
    backspaceRate = "Backspace rate",
    totalKeyPresses = "Total number of key presses",
    active = "Movement rate",
    upright = "Upright rate"
  ))) +
  ylab("Count") +
  xlab("p value") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, family = "serif"),
    strip.text = element_text(size = 14, family = "serif"),
    panel.border = element_rect(fill = NA, color = "grey")
  )
```

```{r}
mean(p_val_df_frag[p_val_df_frag$variable == "active",]$value * 5 * 8 < 0.05)
```



