---
title: "R Notebook"
output: html_notebook
---

```{r setup}
local <- TRUE


dat_dir <- "/Volumes/project/3022000.05/projects/lorkno/data"

# For manuscript images
man_img_dir <- file.path("~/Documents/Writing/Papers/Paper 1/images")

setwd("/Volumes/home/preclineu/lorkno/Documents/clear3-ica")

source("src/load_dependencies.R")
```

# Alex's data

```{r}
dat <- read.table(file.path(dat_dir, "2022-09-14 keystroke detail.psv"), header = TRUE, sep = "|", comment.char = "")
dat
```

```{r}
table(dat$createdOnTimeZone)
```

```{r}
sort(unique(date(dat$createdOnTimestamp / 1000)))
```


# BiAffect data

```{r}
dats_ses[[1]]
```

```{r}
dats_kp[[1]]
```


```{r}
names(dats_ses) <- subjects
biaffect_cors <- bind_rows(dats_ses, .id = "subject") %>%
  mutate(week = week(date)) %>%
  group_by(subject, week) %>%
  select(medianIKD:backspaceRate, totalKeyPresses:bed) %>%
  nest() %>%
  mutate(
    cor_mat = list(cor(data[[1]]))
  )

biaffect_cors
```

```{r}
melt(biaffect_cors$cor_mat[[1]])
```


```{r}
ggplot(melt(biaffect_cors$cor_mat[[102]]), aes(X2, X1)) +
  geom_raster(aes(fill = value)) +
  scale_y_discrete(limits = rev) +
  scale_fill_gradient2(low = muted("blue"), high = muted("red"), mid = "white", 
                       limits = c(-1, 1)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


# Perimenstrual suicidality data

```{r}
all_rep_dat <- read_sav(file.path(dat_dir, "clear3daily_20221205.sav"))
all_rep_dat <- zap_labels(all_rep_dat)
all_rep_dat
```

Write sleep data.

```{r}
write_csv(all_rep_dat %>%
            select(id, daterated, sleepdur_yest, SleepLNQuality),
          file = file.path(dat_dir, "clear3daily_20221205_sleep.csv"))
```


```{r}
table(all_rep_dat$id)
```

```{r warning=FALSE}
max.print <- options("max.print")$max.print
options(max.print = 10 * max.print)

tab1 <- CreateTableOne(data = all_rep_dat)
print(summary(tab1))

options(max.print = max.print)
```

As for the 05-12-2022 snapshot of the data set, we see many spurious days for a couple of subjects (with `daterated` already starting in 1990). Those days contain no actual data, we will simply throw them out of the data set. First select only the columns with relevant data:

```{r}
inclusion <- read_excel("regression_inclusion_20221205.xlsx")$Included
inclusion <- inclusion[1:ncol(all_rep_dat)]
# Relevant self-report data
rel_rep_dat <- all_rep_dat[inclusion == 1]
rel_rep_dat
```


```{r}
all_missing <- data.frame(is.na(rel_rep_dat)) %>%
  rowwise() %>%
  summarize(allMissing = all(c_across(menstrualbleeding:panicattack)))

# Reduced self-report data
red_rep_dat <- rel_rep_dat[!all_missing$allMissing,]
red_rep_dat
```

A reduction of about 22,000 rows. Let's see what TableOne has to say about the data now. But first we throw out the text responses.

```{r warning=FALSE}
max.print <- options("max.print")$max.print
options(max.print = 10 * max.print)

tab1 <- CreateTableOne(data = red_rep_dat)
print(summary(tab1))

options(max.print = max.print)
```

```{r}
complete.cases(red_rep_dat)
```

# Examine all subjects data

Determine data missingness, number of days, etc.

```{r}
dirs <- list.dirs(dat_dir)

pattern <- "sub-([0-9]+)/preproc$"
preproc_paths <- str_subset(dirs, pattern)
subjects <- as.integer(str_match(preproc_paths, pattern)[,2])

n_sub <- length(subjects)
dats_acc <- vector("list", n_sub)
dats_kp <- vector("list", n_sub)
dats_ses <- vector("list", n_sub)
```

First load all relevant data. This might take a while.

```{r}
for (i in 1:length(subjects)) {
  sub <- subjects[i]
  
  print(str_glue("Loading subject {sub}..."))
  
  preproc_path <- preproc_paths[i]
  load(file.path(preproc_path, str_glue("sub-{sub}_preprocessed.rda")))
  
  dats_acc[[i]] <- dat_acc
  dats_kp[[i]] <- dat_kp
  dats_ses[[i]] <- dat_ses
}
```

Save lists separately, using high compression. Hopefully that will reduce loading overhead next time.

```{r}
save(dats_acc, file = file.path(dat_dir, "dats_acc.rda"), compress = "xz")
save(dats_kp, file = file.path(dat_dir, "dats_kp.rda"), compress = "xz")
save(dats_ses, file = file.path(dat_dir, "dats_ses.rda"), compress = "xz")
```

```{r}
load(file.path(dat_dir, "dats_kp.rda"))
load(file.path(dat_dir, "dats_ses.rda"))
```


```{r}
diagnostics <- data.frame(subject = subjects, 
                          n_days = vector("integer", length(subjects)))

for (i in 1:length(subjects)) {
  diagnostics[i,]$n_days <- n_distinct(dats_ses[[i]]$date)
}

diagnostics
```

```{r}
hist(diagnostics$n_days, main = "Histogram of number of key press data days per subject",
     xlab = "Number of days of data per subject")
```

```{r}
all_rep_dat <- read_sav(file.path(dat_dir, "clear3daily_20220801.sav"))
all_rep_dat <- zap_labels(all_rep_dat)
all_rep_dat
```

```{r}
nan_summary <- all_rep_dat %>%
  group_by(id) %>%
  summarize(
    n_days = n(),
    across(menstrualbleeding:LHgroup_perimenstrual_count, ~ mean(is.na(.x)), .names = "{.col}NA")
  )

nan_summary
```

```{r fig.height=20}
par(mfrow = c(10, ceiling(ncol(nan_summary) / 10)))

for (i in 2:ncol(nan_summary)) {
  hist(nan_summary[[i]], main = colnames(nan_summary)[i])
}
```

```{r}
hist(nan_summary[[2]])
```

```{r}
all_rep_dat %>%
  group_by(id) %>%
  summarize(
    n_days = n(),
    na_cases = n() - complete.cases(cur_data())
  )
```



# Subject 2D histogram correlation matrices

```{r}
ls <- get_partitions(dat_dir, date)
subjects <- ls$dat$subjects
dats_kp <- ls$dat$dats_kp
dats_ses <- ls$dat$dats_ses
subject_mats <- ls$subject_mats
```

```{r}
save(subject_mats, file = file.path(dat_dir, "subject_mats.rda"))
```

```{r fig.height=10}
n_sub <- length(subjects)

for (i in 1:n_sub) {
  cors <- cor(sapply(subject_mats[[i]], as.vector))
  g1 <- ggplot(melt(cors), aes(X1, X2)) +
    geom_raster(aes(fill = value)) + 
    scale_y_reverse() +
    ggtitle(str_glue("Subject {subjects[i]}"))
  
  df <- data.frame(mean_cor = colMeans(cors), X1 = 1:ncol(cors))
  g2 <- ggplot(df, aes(X1, mean_cor)) +
    geom_line() +
    geom_point() +
    ylim(0, 1) +
    scale_x_continuous(breaks = seq(0, ncol(cors), 10), 
                       minor_breaks = 1:ncol(cors)) +
    ylab("Mean correlation")
  
  acfs <- apply(cors, 1, acf, lag.max = 300, plot = FALSE)
  df <- melt(t(sapply(acfs, function(x) x$acf)))
  g3 <- ggplot(df, aes(X2, value)) +
    geom_point(alpha = 0.2) +
    scale_x_continuous(breaks = seq(0, ncol(cors), 10), 
                       minor_breaks = 1:ncol(cors)) +
    xlab("X1") +
    ylab("Autocorrelation")
  
  g <- ggarrange(g1, g2, g3, nrow = 3, common.legend = TRUE)
  
  print(g)
  
  ggsave(str_glue("images/autocorrelation/correlations/sub-{subjects[i]}_cor-mat_autocor.png"))
}
```

```{r}
cors <- cor(sapply(subject_mats[[which(subjects == "3083")]], as.vector))
ggplot(melt(cors), aes(X1, X2)) +
  geom_raster(aes(fill = value)) + 
  scale_y_reverse() +
  ggtitle(str_glue("Subject 3083")) +
  labs(fill = "Correlation")

ggsave("images/autocorrelation/correlations/sub-3083_cor_mat.png")
```

# `active` and `totalKeyPresses` histograms

```{r}
active <- lapply(dats_ses, function(df) df$active)
totalKeyPresses <- lapply(dats_ses, function(df) df$totalKeyPresses)

ls <- bin2d(unlist(active), unlist(totalKeyPresses), xBins = 2, yBins = 100)
```

```{r}
ses_hists <- lapply(dats_ses, function(df) {
  df <- df %>%
    group_by(date) %>%
    nest() %>%
    mutate(
      hist = map(data, function(df) {
        binCounts(df$active, 
                  df$totalKeyPresses, 
                  ls$xBreaks, 
                  ls$yBreaks)
      })
    )
  
  df$hist
})
```

```{r}
ls$counts
```


```{r}
counts <- ls$counts %>%
  rownames_to_column("X1") %>%
  melt(id.vars = "X1", variable_name = c("X2")) %>%
  mutate(
    # X1 = case_when(
    #   X1 == "0.245" ~ 0,
    #   TRUE ~ 1
    # ),
    X1 = as.double(X1),
    X2 = as.double(as.character(X2))
  )
  
ggplot(counts, aes(X1, X2)) +
  geom_raster(aes(fill = log(value))) +
  scale_x_continuous(breaks = c(0.25, 0.75), labels = c("FALSE", "TRUE")) +
  ylab("totalKeyPresses") +
  xlab("active")
```

```{r}
df <- dats_ses[[1]] %>%
  group_by(date) %>%
  nest() %>%
  mutate(
    numberOfSessions = nrow(data[[1]]),
    hist = map(data, function(df) {
      binCounts(df$active, 
                df$totalKeyPresses, 
                ls$xBreaks, 
                ls$yBreaks)
    })
  )

df
```

```{r}
df[2,2][[1]]
```

```{r}
ses_hists[[1]][[2]]
```


```{r}
lapply(ses_hists[[1]], function(h) {
  ggplot(melt(h), aes(X2, X1)) +
    geom_raster(aes(fill = value))
})
```

## Bind to 'standard' subject histograms

```{r}
bound_mats <- map2(subject_mats, ses_hists, map2, rbind)
```

```{r}
ggplot(melt(bound_mats[[1]][[1]]), aes(X1, X2)) +
  geom_raster(aes(fill = log(value)))
```

```{r}
save(bound_mats, file = file.path(dat_dir, "bound_mats.rda"))
```


# Confusion matrix of TDIA calculation

```{r}
ls <- parse_tdia_jld(file.path(dat_dir, "tdias_weekly_mats.jld"))
conf_mat <- ls$conf_mat
```

```{r fig.height=8}
# Something to be aware of: X2 indicates the column number, not the row number
ggplot(melt(conf_mat), aes(X2, X1)) +
  geom_raster(aes(fill = value)) + 
  scale_y_reverse() +
  scale_fill_viridis_c() +
  xlab("Predicted subject") +
  ylab("Actual subject") +
  labs(fill = "Count")

# ggsave("images/test_tdia/confusion_matrix.png")

ggplot(melt(conf_mat), aes(X2, X1)) +
  geom_raster(aes(fill = log(value))) + 
  scale_y_reverse()
```

```{r fig.height=8}
b <- barplot(colSums(conf_mat))
axis(1, b, labels = 1:ncol(conf_mat))
```

## Find out what sets apart the attractors

I.e. subjects 5, 23, 38, 39, and 47. First plot data quantity over subjects:

```{r}
dats_kp_alphanum_lens <- lapply(dats_kp, function (dat_kp) {
    nrow(filter(dat_kp, 
                keypress_type == "alphanum", previousKeyType == "alphanum"))
  })
dats_kp_alphanum_lens <- unlist(dats_kp_alphanum_lens)
```

```{r}
tdias_julia_30_blocked_summ <- summarize_tdias(file.path(
                                                 dat_dir, 
                                                 "tdias_flattened.jld"
                                               ), 
                                               n_parts = 30, julia = TRUE)

g_tdia_sub <- plot_tdia_summ(tdias_julia_30_blocked_summ, print_plots = FALSE)[[2]] +
  theme(text = element_text(size = 14))
```

```{r}
g_tdia_sub
```


```{r fig.height=8}
# par(mfrow = c(2, 1))
# 
# b <- barplot(colSums(conf_mat))
# axis(1, b, labels = 1:ncol(conf_mat))
# 
# b <- barplot(dats_kp_alphanum_lens)
# axis(1, b, labels = 1:length(dats_kp_alphanum_lens))

df_conf <- data.frame(subject = 1:ncol(conf_mat), column_count = colSums(conf_mat))
g1 <- ggplot(df_conf, aes(subject, column_count)) +
  geom_col() +
  theme(text = element_text(size=14))

df_kp <- data.frame(subject = 1:ncol(conf_mat), keystroke_count = dats_kp_alphanum_lens)
g2 <- ggplot(df_kp, aes(subject, keystroke_count)) +
  geom_col() +
  theme(text = element_text(size=14))

ggarrange(g1, g2, g_tdia_sub, nrow = 3, align = "hv")

ggsave("images/test_tdia/confusion_matrix_comparison.png")
```

# Cutting away non-baseline data

```{r}
load(file.path(dat_dir, "dat_reg_tdia_30_blocked_expanded_two-handed.rda"))
```

```{r}
start_dates <- read_excel(
  file.path(dat_dir, "CLEAR3_Biaffect_Leow/Metadata/clear3_start_tx_dates.xlsx")) %>%
  mutate(
    ID = as.character(ID),
    study_start_date = as_date(study_start_date),
    treatment_start_date = as_date(treatment_start_date)
  )

start_dates
```

```{r}
min(start_dates$study_start_date)
```


## Regression data

Join start dates with `dat_reg`, then filter out any rows with a date below `study_start_date` and above `treatment_start_date` (if they are present).

```{r}
dat_reg_trimmed <- dat_reg %>%
  left_join(start_dates, c("subject" = "ID")) %>%
  relocate(study_start_date, treatment_start_date, .after = date) %>%
  filter(date >= study_start_date, is.na(treatment_start_date) | date < treatment_start_date) %>%
  ungroup()

dat_reg_trimmed %>%
  filter(subject == "3001", date == ymd("2021-01-02"))
```

```{r}
saveRDS(dat_reg_trimmed, file.path(dat_dir, "dat_reg_trimmed_tdia_30_blocked_expanded_two-handed.rds"))
```

## Raw self-report data

```{r}
dat_rep_trimmed <- all_rep_dat %>%
  mutate(id = as.character(id)) %>%
  left_join(start_dates, c("id" = "ID")) %>%
  relocate(study_start_date, treatment_start_date, .after = daterated) %>%
  filter(daterated >= study_start_date, is.na(treatment_start_date) | daterated < treatment_start_date) %>%
  ungroup()

dat_rep_trimmed
```

```{r fig.height=10}
ggplot(dat_rep_trimmed, aes(daterated, menstrualbleeding)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ id, scales = "free_x")

ggplot(dat_rep_trimmed, aes(daterated, firstdayofperiod)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ id, scales = "free_x")
```

# ICs vs fragmentation

```{r}
load(file.path(dat_dir, "dat_reg_frag_icas.rda"))
na_pat <- readRDS(file.path(dat_dir, "missingness_fragmented.rds"))
```

```{r}
s_df <- data.frame(icas[[1]]$S) %>%
  add_column(subject = subs, date = dates, .before = 1)

s_df
```


```{r}
na_s <- na_pat %>%
  pivot_wider(names_from = modality, values_from = good_block) %>%
  left_join(s_df, by = c("subject", "date")) %>%
  select(!c(X3, X4, X5)) %>%
  pivot_longer(X1:X2, names_to = "ic")

na_s
```

```{r fig.height=15}
ggplot(na_s, aes(date)) +
  geom_line(aes(y = value, color = ic)) +
  geom_tile(aes(y = -3, fill = self_report)) +
  scale_fill_manual(values = c("grey", "steelblue")) +
  facet_wrap(~ subject, scales = "free_x")
```

# IC 3 vs participant means

`prep_reg_dat` can be found in `multilevel_modelling.Rmd`.

```{r}
ls <- prep_reg_dat(
  dat_path = file.path(dat_dir, "dat_reg_fragmented_no-tdia.rds"),
  ica_path = file.path(dat_dir, "dat_reg_frag_icas.rda")
)

dat_reg_no_tdia <- ls$dat_reg
dat_bi <- ls$dat_bi
dat_sr <- ls$dat_sr
icas <- ls$icas
subs <- ls$subs
dates <- ls$dates
dat_c <- ls$dats_c[[1]]
```

```{r}
ic_df <- data.frame(icas[[1]]$S) %>%
  add_column(subject = subs, date = dates)

dat_ica_sr <- dat_reg_no_tdia %>%
  filter(modality == "self_report") %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  inner_join(ic_df, by = c("subject", "date"))

dat_ica_sr
```

```{r}
all_means <- dat_ica_sr %>%
  group_by(subject) %>%
  summarize(across(c(ASIQ9_wishdead:mastery, X3), mean)) %>%
  rowwise(subject, X3) %>%
  summarize(mean = mean(c_across(ASIQ9_wishdead:mastery)))

all_means
```

```{r}
ggplot(all_means, aes(X3, mean)) +
  geom_point()
```

```{r}
cor.test(all_means$X3, all_means$mean)
```

```{r}
time_mean <- dat_ica_sr %>%
  rowwise(subject, date, X3) %>%
  summarize(
    self_report = mean(c_across(ASIQ9_wishdead:mastery)),
    sd = sd(c_across(ASIQ9_wishdead:mastery))
  ) %>%
  pivot_longer(c(X3, self_report), names_to = "modality")

time_mean
```

```{r fig.height=10}
ggplot(time_mean, aes(date, value, color = modality)) +
  geom_line() +
  labs(x = "Date", y = "Value", color = "Modality") +
  scale_x_date(breaks = NULL) +
  scale_color_hue(labels = c("Mean self-report rating", "Independent component 3")) +
  facet_wrap(~ subject, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8))
```

All in all, it is quite clear that IC 3 corresponds to the inverted mean item rating for several subjects.

```{r}
ggplot(time_mean %>% filter(subject == "3083"), aes(date, value, color = modality)) +
  geom_line(lwd = 1) +
  scale_color_manual(values = c("lightcoral", "turquoise4"),
                     labels = c("Mean self-report rating", "Independent component 3")) +
  labs(x = "Date", y = "Value", color = "Modality") +
  theme_minimal()

ggsave(file.path(man_img_dir, "ic3_timecourse.pdf"))
```

```{r}
n_distinct(filter(time_mean, subject == "3083")$date)
```

# Key press bubble plot

```{r}
ls <- load_subject_key_data(dat_dir)
dats_kp <- ls$dats_kp
dats_ses <- ls$dats_ses
subjects <- ls$subjects
```

## Intermezzo to port some data to parquet files

```{r}
library(arrow)

names(dats_ses) <- subjects
dat_ses <- bind_rows(dats_ses, .id = "subject") %>%
  arrow_table() %>%
  write_parquet(file.path(dat_dir, "dat_ses.parquet"))
```

```{r}
dats_kp[subjects == "3029"][[1]] %>% 
  select(createdOnTimestamp, timestamp, sessionNumber, createdOnTimeZone) %>%
  mutate(
    createdOnTimestamp = round(createdOnTimestamp / 1e3),
    first_stroke = c(1, diff(sessionNumber))
  ) %>%
  filter(first_stroke == 1) %>%
  mutate(timestampLocal = timestamp + createdOnTimeZone * 36, .before = 3)
```

## End of intermezzo

```{r}
sizes <- sapply(dats_ses, nrow)

dat_ses <- dats_ses[sizes == max(sizes)][[1]]
```

```{r}
dat_ses %>%
  group_by(date) %>%
  summarize(n = n())
```

```{r}
hist(1 / (dat_ses$medianIKD[dat_ses$medianIKD < 0.5]))
```


```{r fig.height=10}
reverse2_trans <- function() {
  trans_new(
    "reverse2",
    function(x) as.Date(-1 * as.numeric(x), ymd("1970-01-01")), # Force values to be numeric for Date objects
    function(x) as.Date(-1 * as.numeric(x), ymd("1970-01-01"))
  )
}

ggplot(dat_ses %>%
         filter(date > ymd("2020-12-31"), date < ymd("2021-12-31")), 
       aes(hour, date, size = totalKeyPresses, color = medianIKD)) +
  geom_point() +
  xlab("Hour") +
  ylab("Date") +
  scale_y_continuous(
    trans = c("date", "reverse2")
  ) +
  scale_size_continuous("Number of key presses") +
  scale_color_viridis_c("Delay between keys (s)", limits = c(0, 0.5)) +
  theme(text = element_text(size = 18))
```

```{r}
ggplot(dat_ses %>%
         group_by(hour) %>%
         summarize(
           sd = sd(medianIKD, na.rm = TRUE),
           medianIKD = mean(medianIKD, na.rm = TRUE)
         ), 
       aes(hour, medianIKD)) +
  geom_line(group = 1) +
  geom_errorbar(aes(ymin = medianIKD - sd, ymax = medianIKD + sd))
```

# Modality temporal alignment

```{r}
load(file.path(dat_dir, "sub-3009/preproc/sub-3009_preprocessed.rda"))
```

```{r}
acc_path <- file.path(dat_dir, "sub-3009/raw/sub-3009_accelerometer.psv")
raw_acc <- read.table(acc_path, header = TRUE, sep = "|", strip.white = TRUE, fill = TRUE)

dat_acc <- raw_acc %>%
  mutate(
    # Timezone is coded as number of hours deviating from UCT times 100
    # (e.g. -600), so multiply by 36 to get offset in seconds.
    sessionTimestampLocal = as_datetime(sessionTimestamp + createdOnTimeZone * 36),
    sessionNumber = cumsum(sampleNumber == 1)
  ) %>%
  rename(x = xCoord, y = yCoord, z = zCoord)
```

```{r}
saveRDS(dat_acc, file.path(dat_dir, "sub-3009/preproc/sub-3009_semi-raw-dat-acc.rds"))
```

```{r}
dat_acc <- readRDS(file.path(dat_dir, "sub-3009/preproc/sub-3009_semi-raw-dat-acc.rds"))
```


ymd("2021-02-28")

```{r}
timestamps_kp <- dat_kp %>%
  select(c(keypressTimestampLocal)) %>%
  rename(timestamp = keypressTimestampLocal) %>%
  add_column(modality = "Keyboard") %>%
  filter(date(timestamp) == ymd("2021-02-28"))

timestamps_acc <- dat_acc %>%
  filter(date(sessionTimestampLocal) == ymd("2021-02-28")) %>%
  mutate(timestamp = sessionTimestampLocal + seconds((sampleNumber - 1) * 0.1)) %>%
  select(c(timestamp)) %>%
  add_column(modality = "Accelerometer")

timestamps_sr <- data.frame(timestamp = parse_date_time("2021-02-28 19:30:15", "Y-m-d H:M:S"),
                            modality = "Self-report")

timestamps <- bind_rows(timestamps_kp, timestamps_acc, timestamps_sr) %>%
  filter(timestamp >= parse_date_time("2021-02-28 19:30", "Y-m-d H:M"),
         timestamp < parse_date_time("2021-02-28 19:33", "Y-m-d H:M"))
```

```{r fig.height=20, fig.width=20}
ggplot(timestamps, aes(timestamp, modality)) +
  geom_point(aes(color = modality), size = 8) +
  labs(x = "Time (seconds)", y = "Modality") +
  guides(color = "none") +
  theme_minimal() +
  theme(
    text = element_text(size = 32),
    axis.text = element_text(size = 35)
  )

blog_dir <- "~/Documents/Writing/Papers/Paper 1/Blog post"

ggsave(file.path(blog_dir, "Figure1_sampling_discrepancies.png"))
```

```{r}
timestamps_kp <- dat_kp %>%
  select(c(keypressTimestampLocal)) %>%
  rename(timestamp = keypressTimestampLocal) %>%
  add_column(modality = "Keyboard") %>%
  filter(date(timestamp) >= ymd("2021-02-25"), date(timestamp) <= ymd("2021-02-28"))

timestamps_acc <- dat_acc %>%
  filter(date(sessionTimestampLocal) >= ymd("2021-02-25"), date(sessionTimestampLocal) <= ymd("2021-02-28")) %>%
  mutate(timestamp = sessionTimestampLocal + seconds((sampleNumber - 1) * 0.1)) %>%
  select(c(timestamp)) %>%
  add_column(modality = "Accelerometer")

dates <- unique(date(timestamps_kp$timestamp))
timestamps_sr <- data.frame(timestamp = dates + hours(19),
                            modality = "Self-report")

timestamps <- bind_rows(timestamps_kp, timestamps_acc, timestamps_sr)
```

```{r fig.height=10, fig.width=10}
ggplot(timestamps, aes(timestamp, modality)) +
  geom_point(aes(color = modality), size = 3) +
  labs(x = "Time", y = "Modality") +
  guides(color = "none") +
  theme_minimal() +
  theme(
    text = element_text(size = 18),
    axis.text = element_text(size = 20)
  )
```

```{r fig.height=10, fig.width=10}
df <- data.frame(
  timestamp = rep(c(ymd("2021-02-25"), 
                    ymd("2021-02-26"), 
                    ymd("2021-02-27"), 
                    ymd("2021-02-28")), 
                  3),
  modality = rep(c("Accelerometer", "Keyboard", "Self-report"), each = 4)
)

ggplot(df, aes(timestamp, modality)) +
  geom_point(aes(color = modality), size = 3) +
  labs(x = "Time", y = "Modality") +
  guides(color = "none") +
  theme_minimal() +
  theme(
    text = element_text(size = 18),
    axis.text = element_text(size = 20)
  )
```

# Self-report dimensionality

```{r}
all_rep_dat <- read_sav(file.path(dat_dir, "clear3daily_20221205.sav"))
all_rep_dat <- zap_labels(all_rep_dat)
all_rep_dat
```

```{r}
all_rep_dat %>%
  filter(daterated == max(daterated))
```


```{r}
rep_dat <- readRDS(file.path(dat_dir, "dat_reg_fragmented_no-tdia.rds")) %>% 
  filter(modality == "self_report") %>%
  select(!c(modality, study_start_date, treatment_start_date))

rep_dat
```

```{r fig.height=15, fig.width=15}
ggplot(rep_dat %>% filter(subject == 3009), aes(date, value)) +
  geom_line() +
  facet_wrap(~ variable, ncol = 1, strip.position = "left") +
  xlab("Date") +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.text = element_blank(),
    # strip.clip = "off",
    strip.background = element_blank(),
    # strip.text = element_text(size = 12),
    axis.text.x = element_text(size = 23),
    axis.title.x = element_text(size = 25)
  )

ggsave(file.path(blog_dir, "Figure2_self-report_dimensionality.png"))
```


```{r}
sr_example <- all_rep_dat %>%
  select(c(id, daterated, sleepdur_yest:mastery)) %>%
  filter(id == 3009) %>%
  pivot_longer(sleepdur_yest:mastery, names_to = "variable") %>%
  mutate(variable = as.numeric(factor(variable)))

sr_example
```

```{r fig.height=15, fig.width=15}
ggplot(sr_example, aes(daterated, value)) +
  geom_line() +
  facet_wrap(~ variable, ncol = 1, strip.position = "left") +
  xlab("Date") +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.text = element_blank(),
    # strip.clip = "off",
    strip.background = element_blank(),
    # strip.text = element_text(size = 12),
    axis.text.x = element_text(size = 18),
    axis.title.x = element_text(size = 20)
  )
```

