---
title: "R Notebook"
output: html_notebook
---

```{r setup}
local <- TRUE

wd <- ""

if (!local) {
  # Cluster environment
  wd <- "~/Documents/Autocorrelation"
  # Base directory for the data set we will create
  dat_dir <- "/project_cephfs/3022017.02/projects/lorkno/data"
} else {
  wd <- "/Volumes/home/preclineu/lorkno/Documents/Autocorrelation"
  # Base directory for the data set we will create
  dat_dir <- "~/HPC_project/data"
}

model_dir <- file.path(dat_dir, "complete_data_models")

# For manuscript images
man_img_dir <- file.path("~/Documents/Writing/Papers/Paper 1/images")

# For some reason this sets the pwd only for the current scope, which means it 
# does not affect anything outside an if block if you set it there.
# So that's why it's here instead of above.
setwd(wd) 

source("src/load_dependencies.R")
library(fastICA)
```

# Get/set seed

I would like to avoid setting the seed manually. I'll record my seed from today (14-06-2023) and use that in later runs of this script.

```{r}
# if(!exists(".Random.seed")) set.seed(NULL)
# seed <- .Random.seed
# saveRDS(seed, file.path(dat_dir, "sens_seed.rds"))

seed <- readRDS(file.path(dat_dir, "sens_seed.rds"))
```

# Prepare ICAs

For now, only do the 5-component solution.

```{r}
replicate_icas <- function(dat_path, n = 100, out_path = NULL, sub_norm = FALSE) {
  set.seed(seed)
  
  ### Data prep
  
  dat_reg <- readRDS(dat_path)
  
  dat_reg_sr <- dat_reg %>%
    filter(modality == "self_report", !is.na(value)) %>%
    select(!c(study_start_date, treatment_start_date, modality)) %>%
    group_by(variable) %>%
    mutate(
      logged = case_when(
        min(value) == 0 ~ log(value + 1), # To prevent log(0) = NA values
        TRUE ~ log(value)
      ),
      scaled = scale(value)
    ) %>%
    ungroup()
  
  if (sub_norm) {
    # Normalize variance within subjects
    dat_reg_sr <- dat_reg_sr %>%
      group_by(subject, variable) %>%
      mutate(logged = scale(logged, scale = FALSE)) %>%
      ungroup()
  }
  
  df <- dat_reg_sr %>%
    select(!c(value, scaled)) %>% 
    pivot_wider(names_from = variable, values_from = logged) %>%
    arrange(subject, date) 
  
  subs <- df$subject
  dates <- df$date
  
  df <- df %>%
    select(!c(subject, date))
  
  # t x q
  mat <- sapply(df, as.vector)
  
  means <- colMeans(mat)
  
  ### Running the ICAs
  
  icas <- replicate(
    n = n, 
    fastICA(mat, n.comp = 5, alg.typ = "parallel", fun = "logcosh"),
    simplify = FALSE
  )
  
  ls <- list(icas = icas, subs = subs, dates = dates)
  
  if (!is.null(out_path))
    saveRDS(ls, file = out_path)
  
  ls
}

prep_reg_dat <- function(dat_path, icas, subs, dates, general_idx) {
  # Semi-contiguous or fragmented data
  dat_reg <- readRDS(dat_path)
  
  dat_bi <- dat_reg %>%
    filter(modality == "biaffect") %>%
    select(!c(study_start_date, treatment_start_date, modality)) %>%
    pivot_wider(names_from = variable, values_from = value) %>%
    select(-any_of(c("partition_idx")))
  
  # Select components that correspond to general affect components
  dats <- map2(icas, general_idx, function(ica, idx) {
    data.frame(X = ica$S[,idx]) %>%
      add_column(subject = subs, date = dates) %>%
      inner_join(dat_bi, by = c("subject", "date")) %>%
      mutate(
        week = week(date),
        sub_week = paste(subject, week, sep = " - ") # For convenience
      ) %>%
      relocate(subject, date, week, sub_week, .before = X)
  })
  
  dats_f <- lapply(dats, function(dat) {
    dat %>%
      group_by(sub_week) %>%
      filter(n() > 1) %>%
      ungroup()
  })
  
  dats_c <- lapply(dats_f, function(dat) {
    dat %>%
      mutate(
        totalKeyPresses = log(totalKeyPresses),
        across(medianIKD:bed, scale)
      )
  })
  
  dats_c
}
```

```{r}
ls <- replicate_icas(file.path(dat_dir, "dat_reg_semi_contiguous_no-tdia.rds"),
                     out_path = file.path(dat_dir, "icas_sens_no-tdia.rds"))

icas <- ls$icas
subs <- ls$subs
dates <- ls$dates
```

We use `sens_helper.R` to mark the general affect components and save the marked component indices to `general_idx_sens.rds`.

```{r}
general_idx <- readRDS(file.path(dat_dir, "general_idx_sens.rds"))
general_idx <- as.integer(general_idx)
```

# Run models

Prepare data.

```{r}
dats_c <- prep_reg_dat(file.path(dat_dir, "dat_reg_semi_contiguous_no-tdia.rds"),
                       icas, subs, dates,
                       general_idx)
```

Create models.

```{r}
models <- lapply(dats_c, function(dat_c) {
  lme(X ~ medianIKD +
        percent95IKD +
        madIKD +
        autocorrectRate +
        backspaceRate +
        totalKeyPresses +
        active +
        upright,
      random = ~ 1 | subject / week,
      data = dat_c)
})
```

Extract p values.

```{r}
p_vals <- sapply(models, function(m) {
  summary(m)$tTable[,5]
})

p_val_df <- data.frame(t(p_vals)) %>%
  pivot_longer(everything(), names_to = "variable") %>%
  filter(variable != "X.Intercept.")

p_val_df
```


Plot p values.

```{r fig.height=10}
ggplot(p_val_df, aes(value)) +
  geom_histogram(bins = 60) +
  facet_wrap(~ variable, scales = "free_x", labeller = as_labeller(c(
    medianIKD = "Median IKD",
    percent95IKD = "95th percentile IKD",
    madIKD = "MAD IKD",
    autocorrectRate = "Autocorrect rate",
    backspaceRate = "Backspace rate",
    totalKeyPresses = "Total number of key presses",
    active = "Movement rate",
    upright = "Upright rate"
  ))) +
  ylab("Count") +
  xlab("p value") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, family = "serif"),
    strip.text = element_text(size = 14, family = "serif"),
    panel.border = element_rect(fill = NA, color = "grey")
  )
```

```{r fig.height=10}
ggplot(p_val_df, aes(value)) +
  geom_histogram(bins = 60) +
  facet_wrap(~ variable, scales = "free_x", labeller = as_labeller(c(
    medianIKD = "Median IKD",
    percent95IKD = "95th percentile IKD",
    madIKD = "MAD IKD",
    autocorrectRate = "Autocorrect rate",
    backspaceRate = "Backspace rate",
    totalKeyPresses = "Total number of key presses",
    active = "Movement rate",
    upright = "Upright rate"
  ))) +
  ylab("Count") +
  xlab("p value") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, family = "serif"),
    strip.text = element_text(size = 14, family = "serif"),
    panel.border = element_rect(fill = NA, color = "grey")
  )
```

Significant p values.

```{r}
bonf_p_vals <- p_val_df[p_val_df$variable == "active",]$value * 5 * 8
mean(bonf_p_vals < 0.05)
```

```{r}
range(bonf_p_vals)
table(signif(bonf_p_vals, 2))
mean(bonf_p_vals > 0.001)
```

```{r}
round(bonf_p_vals, 5)
```


Proportion of smallest p values.

```{r}
mean(p_val_df[p_val_df$variable == "active",]$value < 1e-5)
```

Beta values.

```{r}
betas <- sapply(models, function(m) summary(m)$tTable["active", 1])
range(betas)
plot(bonf_p_vals, betas)
```

```{r}
table(signif(betas, 2))
```


Does the size of the p values coincide with the component splitting for the non-fragmented ICA? (Conclusion: Yes, it does.)

```{r fig.height=10}
mix_dfs <- lapply(icas, function(ica) {
  data.frame(ica$A) %>%
    rename_with(function(x) mat_names) %>%
    add_column(ic = 1:nrow(ica$A), .before = 1) %>%
    pivot_longer(!ic, names_to = "variable", values_to = "loading") %>%
    mutate(
      variable = var_dict[variable],
      variable = fct_rev(factor(variable, var_dict, ordered = TRUE)),
      variable_code = as.numeric(variable)
    )
})

big_mix_df <- bind_rows(mix_dfs, .id = "run") %>%
  mutate(run = as.integer(run))

ps <- p_vals["active",]

for (i in seq(5, 100, 5)) {
  p_dict <- sprintf("%d: %#.2g", (i-4):i, ps[(i-4):i])
  names(p_dict) <- (i-4):i
  
  g <- ggplot(big_mix_df %>% filter((i - 5) < run, run <= i), 
         aes(loading, variable, fill = loading)) +
    geom_col() +
    facet_grid(vars(run), vars(ic), labeller = labeller(run = p_dict)) +
    scale_fill_gradient2(low = muted("blue"), mid = "white", high = muted("red"), 
                         name = "Loading") +
    xlim(-0.6, 0.6) +
    xlab("Loading") +
    ylab("Variable") +
    theme(text = element_text(size = 22, family = "serif"),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          axis.text.y = element_text(size = 5))
  
  print(g)
}
```

# Fragmented data

```{r}
ls <- replicate_icas(file.path(dat_dir, "dat_reg_fragmented_no-tdia.rds"))
icas_frag <- ls$icas
subs_frag <- ls$subs
dates_frag <- ls$dates
```

```{r fig.height=10}
fancy_mix_fig(icas_frag[[1]])
```

Use sens_helper.R to get general_idx_frag.

```{r}
general_idx_frag <- readRDS(file.path(dat_dir, "general_idx_frag_sens2.rds"))
general_idx_frag <- as.integer(general_idx_frag)

dats_c_frag <- prep_reg_dat(file.path(dat_dir, "dat_reg_fragmented_no-tdia.rds"),
                       icas_frag, subs_frag, dates_frag,
                       general_idx_frag)

models_frag <- lapply(dats_c_frag, function(dat_c) {
  lme(X ~ medianIKD +
        percent95IKD +
        madIKD +
        autocorrectRate +
        backspaceRate +
        totalKeyPresses +
        active +
        upright,
      random = ~ 1 | subject / week,
      data = dat_c)
})

p_vals_frag <- sapply(models_frag, function(m) {
  summary(m)$tTable[,5]
})

p_val_df_frag <- data.frame(t(p_vals_frag)) %>%
  pivot_longer(everything(), names_to = "variable") %>%
  filter(variable != "X.Intercept.")
```

```{r fig.height=10}
ggplot(p_val_df_frag, aes(value)) +
  geom_histogram(bins = 60) +
  facet_wrap(~ variable, scales = "free_x", labeller = as_labeller(c(
    medianIKD = "Median IKD",
    percent95IKD = "95th percentile IKD",
    madIKD = "MAD IKD",
    autocorrectRate = "Autocorrect rate",
    backspaceRate = "Backspace rate",
    totalKeyPresses = "Total number of key presses",
    active = "Movement rate",
    upright = "Upright rate"
  ))) +
  ylab("Count") +
  xlab("p value") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, family = "serif"),
    strip.text = element_text(size = 14, family = "serif"),
    panel.border = element_rect(fill = NA, color = "grey")
  )

ggplot(p_val_df_frag %>% filter(variable == "active"), aes(value)) +
  geom_histogram(bins = 60) +
  ylab("Count") +
  xlab("p value") +
  labs(title = "Movement rate significance") +
  theme_minimal() +
  theme(
    text = element_text(size = 20, family = "serif"),
    title = element_text(size = 24, family = "serif"),
    panel.border = element_rect(fill = NA, color = "grey")
  )
```

Significant p values.

```{r}
bonf_p_vals <- p_val_df_frag[p_val_df_frag$variable == "active",]$value * 5 * 8
mean(bonf_p_vals < 0.05)
```

```{r}
range(bonf_p_vals)
table(signif(bonf_p_vals, 2))
mean(bonf_p_vals > 0.001)
```

Proportion of smallest p values.

```{r}
mean(p_val_df_frag[p_val_df_frag$variable == "active",]$value < 1e-5)
```

Beta values.

```{r}
table(signif(betas, 2))
```

```{r}
betas <- sapply(models_frag, function(m) summary(m)$tTable["active", 1])
range(betas)
plot(bonf_p_vals, betas)
```


Plot mixing matrices with their p values to check if the higher p values correspond to the split ICA solutions (conclusion: They do).

```{r}
mix_dfs <- lapply(icas_frag, function(ica) {
  data.frame(ica$A) %>%
    rename_with(function(x) mat_names) %>%
    add_column(ic = 1:nrow(ica$A), .before = 1) %>%
    pivot_longer(!ic, names_to = "variable", values_to = "loading") %>%
    mutate(
      variable = var_dict[variable],
      variable = fct_rev(factor(variable, var_dict, ordered = TRUE)),
      variable_code = as.numeric(variable)
    )
})

big_mix_df <- bind_rows(mix_dfs, .id = "run") %>%
  mutate(run = as.integer(run))
  
big_mix_df
```

```{r fig.height=10}
p_vals <- p_vals_frag["active",]

for (i in seq(5, 100, 5)) {
  p_dict <- sprintf("%d: %#.2g", (i-4):i, p_vals[(i-4):i])
  names(p_dict) <- (i-4):i
  
  g <- ggplot(big_mix_df %>% filter((i - 5) < run, run <= i), 
         aes(loading, variable, fill = loading)) +
    geom_col() +
    facet_grid(vars(run), vars(ic), labeller = labeller(run = p_dict)) +
    scale_fill_gradient2(low = muted("blue"), mid = "white", high = muted("red"), 
                         name = "Loading") +
    xlim(-0.6, 0.6) +
    xlab("Loading") +
    ylab("Variable") +
    theme(text = element_text(size = 22, family = "serif"),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          axis.text.y = element_text(size = 5))
  
  print(g)
}
```

Plot a mixing matrix with the general component and the lethargy component next to each other.

```{r fig.height=10}
library(ggh4x)

cols <- c("lightcoral", "aquamarine3", "turquoise4", "sienna", "darkorange3", 
          "tan1", "indianred3", "steelblue")

n_ics <- 5

# Custom annotation function
ann <- function(grob,
                xmin = -Inf, xmax = Inf,
                ymin = -Inf, ymax = Inf,
                data) {
  layer(data = data, stat = StatIdentity, position = PositionIdentity,
        geom = ggplot2:::GeomCustomAnn,
        inherit.aes = TRUE, params = list(grob = grob,
                                          xmin = xmin, xmax = xmax,
                                          ymin = ymin, ymax = ymax))
}

A <- icas_frag[[6]]$A
mix_df_general <- data.frame(A) %>%
  rename_with(function(x) mat_names) %>%
  add_column(ic = paste("IC", 1:nrow(A)), .before = 1) %>%
  pivot_longer(!ic, names_to = "variable", values_to = "loading") %>%
  mutate( # Add some questionnaire grouping information.
    questionnaire = case_when(
      str_starts(variable, "ASIQ") ~ "ASIQ",
      variable == "belonging_" ~ "INQ",
      str_starts(variable, "BAM") ~ "BAM",
      str_starts(variable, "BITe") ~ "BITe",
      str_starts(variable, "DRSP") ~ "DRSP",
      variable == "mastery" | variable == "wishsleep" | variable == "wantedkill" ~ "Misc",
      variable == "PANAS_happy"  ~ "PANAS"
    ),
    variable = var_dict[variable],
    variable = fct_rev(factor(variable, var_dict, ordered = TRUE)),
    variable_code = as.numeric(variable),
    class = "General affect"
  )

A <- icas_frag[[11]]$A
mix_df_lethargy <- data.frame(A) %>%
  rename_with(function(x) mat_names) %>%
  add_column(ic = paste("IC", 1:nrow(A)), .before = 1) %>%
  pivot_longer(!ic, names_to = "variable", values_to = "loading") %>%
  mutate( # Add some questionnaire grouping information.
    questionnaire = case_when(
      str_starts(variable, "ASIQ") ~ "ASIQ",
      variable == "belonging_" ~ "INQ",
      str_starts(variable, "BAM") ~ "BAM",
      str_starts(variable, "BITe") ~ "BITe",
      str_starts(variable, "DRSP") ~ "DRSP",
      variable == "mastery" | variable == "wishsleep" | variable == "wantedkill" ~ "Misc",
      variable == "PANAS_happy"  ~ "PANAS"
    ),
    variable = var_dict[variable],
    variable = fct_rev(factor(variable, var_dict, ordered = TRUE)),
    variable_code = as.numeric(variable),
    class = "Lethargy"
  )

mix_df <- rbind(mix_df_general, mix_df_lethargy)

separator_df <- filter(mix_df, class == "General affect", ic == "IC 5")

ggplot(mix_df, aes(loading, 
                   variable_code, 
                   fill = questionnaire, 
                   alpha = abs(loading))) +
  # annotate(geom = "rect", xmin = -0.5, xmax = 0.5, ymin = 0.5, ymax = 9.5, fill = alpha("steelblue", 0.5)) + 
  # annotate(geom = "tile", x = 0, y = 11, width = 1, height = 3, fill = alpha("sienna", 0.5)) + 
  geom_col(aes(group = questionnaire), orientation = "y") +
  # geom_vline(aes(xintercept = x), data = separator_df) +
  ann(linesGrob(gp = gpar(col = "grey50", lwd = 1.5)), xmin = 0.71, xmax = 0.71, ymin = -1, ymax = 37.5, data = separator_df) +
  facet_nested(~ class + ic) +
  scale_y_continuous(breaks = 1:max(mix_df$variable_code), 
                     labels = levels(mix_df$variable)) +
  scale_fill_manual(values = cols, name = "Questionnaire") +
  scale_alpha_continuous(guide = "none") +
  xlim(-0.6, 0.6) +
  xlab("Loading") +
  ylab("Variable") +
  coord_cartesian(clip = "off") +
  theme(text = element_text(size = 22),
        axis.text.x = element_text(size = 14, angle = 90, vjust = 0.5, hjust = 1),
        # axis.text.y = element_blank(),
        axis.text.y = element_text(size = 10, margin = margin(0, 0, 0, 0)),
        axis.ticks.y = element_blank(),
        plot.margin = unit(c(1, 1, 1, 1), "lines"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgrey"),
        strip.background = element_rect("white"),
        ggh4x.facet.nestline = element_line(color = "grey50"))

ggsave(file.path(man_img_dir, "mixing_matrix_classes_frag.pdf"),
       width = 15, height = 10)
```


# Mean-centered

```{r fig.height=10}
ls <- replicate_icas(file.path(dat_dir, "dat_reg_fragmented.rds"), 
                     sub_norm = TRUE)
icas_cent <- ls$icas
subs_cent <- ls$subs
dates_cent <- ls$dates

fancy_mix_fig(icas_cent[[2]], 
              save_path = file.path(man_img_dir, "mixing_matrix_mean_centred.pdf"))
```

Use `sens_helper.R` to get `general_idx_cent`.

```{r}
general_idx_cent <- readRDS(file.path(dat_dir, "general_idx_cent.rds"))
general_idx_cent <- as.integer(general_idx_cent)

dats_c_cent <- prep_reg_dat(file.path(dat_dir, "dat_reg_fragmented.rds"),
                            icas_cent, subs_cent, dates_cent,
                            general_idx_cent)

models_cent <- lapply(dats_c_cent, function(dat_c) {
  lme(X ~ medianIKD +
        percent95IKD +
        madIKD +
        autocorrectRate +
        backspaceRate +
        totalKeyPresses +
        active +
        upright,
      random = ~ 1 | subject / week,
      data = dat_c)
})

p_vals_cent <- sapply(models_cent, function(m) {
  summary(m)$tTable[,5]
})

p_val_df_cent <- data.frame(t(p_vals_cent)) %>%
  pivot_longer(everything(), names_to = "variable") %>%
  filter(variable != "X.Intercept.")
```

```{r fig.height=10}
ggplot(p_val_df_cent, aes(value)) +
  geom_histogram(bins = 60) +
  facet_wrap(~ variable, scales = "free_x", labeller = as_labeller(c(
    medianIKD = "Median IKD",
    percent95IKD = "95th percentile IKD",
    madIKD = "MAD IKD",
    autocorrectRate = "Autocorrect rate",
    backspaceRate = "Backspace rate",
    totalKeyPresses = "Total number of key presses",
    active = "Movement rate",
    upright = "Upright rate"
  ))) +
  ylab("Count") +
  xlab("p value") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, family = "serif"),
    strip.text = element_text(size = 14, family = "serif"),
    panel.border = element_rect(fill = NA, color = "grey")
  )
```

```{r}
mean(p_val_df_cent[p_val_df_cent$variable == "active",]$value * 5 * 8 < 0.05)
```

```{r}
dat_reg_frag <- readRDS(file.path(dat_dir, "dat_reg_fragmented.rds"))
ica <- icas_cent[[2]]
  
dat_bi_frag <- dat_reg_frag %>%
  filter(modality == "biaffect") %>%
  select(!c(study_start_date, treatment_start_date, modality)) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  select(!partition_idx)

dat_joined_cent <- data.frame(ica$S) %>%
  add_column(subject = subs_cent, date = dates_cent) %>%
  inner_join(dat_bi_frag, by = c("subject", "date")) %>%
  mutate(
    week = week(date),
    sub_week = paste(subject, week, sep = " - ") # For convenience
  ) %>%
  relocate(subject, date, week, sub_week, .before = X1) %>%
  group_by(sub_week) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  mutate(
    totalKeyPresses = log(totalKeyPresses),
    across(medianIKD:mean_accuracy, scale)
  )

dat_joined_cent
```

```{r}
m5.cent.X1.nlme <- lme(X1 ~ medianIKD +
                         percent95IKD +
                         madIKD +
                         autocorrectRate +
                         backspaceRate +
                         totalKeyPresses +
                         active +
                         upright,
                       random = ~ 1 | subject / week,
                       data = dat_joined_cent)

m5.cent.all.nlme <- dat_joined_cent %>%
  summarize(across(starts_with("X", ignore.case = FALSE), 
                   ~ list(update(m5.cent.X1.nlme, 
                                 reformulate(".", cur_column())))))
```

```{r paged.print=FALSE}
lapply(m5.cent.all.nlme, function(m) summary(m[[1]]))
```

